### 2021-03-25 analysis/xor-some-prolific-pilot1.Rmd
minor fix

---
title: "XOR-Some Prolific Pilot 1"
author: "Polina Tsvilodub"
date: "3/24/2021"
output: github_document
---

In this first pilot on Prolific we test the implementation of the xor study in magpie. The structure of the experiment was as follows:
Participants read instructions, completed three example trials, and then completed 8 main blocks consisting of 4 xor and 4 some items. Each main block had the following structure: Participants read the background story, answered one comprehension question, then answered competence / relevance / prior questions in randomized order; then they read another 3 comprehension questions, after which the critical utterance was added below the background story. They answered the inference strength question, and then competence / relevance questions in randomized order again.

N=10 participants were recruited for this pilot and compensated 2 pounds/participant.

```{r libraries}
library(tidyverse)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r anonymization, include=FALSE}
#d_raw <- read_csv("~/Documents/Research/XOR/results_58_xor-some-Prolific-pilot1_N10.csv")
# remove IDs
#d_raw %>% select(-prolific_id) %>% write_csv("../data/pilots/results_58_xor-some-Prolific-pilot1_N10.csv") 
```

```{r data}
d <- read_csv("../data/pilots/results_58_xor-some-Prolific-pilot1_N10.csv")
```

Checking if there are any comments indicating technical issues:
```{r comments}
d %>% distinct(comments)
```

Check native languages. In the main experiment, participants not indicating English as (one of) their native language(s) will be excluded. We also compute some participant demographics.
```{r languages}
d %>% distinct(languages)
# exclude non-natives if necessary
d_native <- d

d_native %>% pull(age) %>% mean(., na.rm = T) 
d_native %>% count(gender) %>% mutate(n = n/88)
```

Next, we check whether all the conditions were used correctly.

```{r counts}
# check xor/some vs. trial type
d_native %>% count(main_type, condition) 
# check xor/some vs. experimental condition
d_native %>% count(main_type, exp_condition)
# count items used
d_native %>% count(title)
```

Next, we would exclude participants based on their ratings in the main / example trials (and possibly the bot check) according to preregistered exclusion criteria. For now, all the data is considered.
```{r clean}
d_native %>% group_by(submission_id) %>% count(response)
d_main <- d_native %>% filter(trial_name != "example")
d_exmpl <- d_native %>% filter(trial_name == "example")
d_critical <- d_main %>% filter(condition == "critical")

# get overall mean ratings / subject
d_native %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)
```
Plot responses on example questions by question type

Plot test questions by type
```{r}
d_test <- d_main %>% rowwise() %>% filter(condition == "test") %>% 
  mutate(test_condition = substr(test_question, 6, 9),
         test_condition = ifelse(test_condition == "fals", "false", 
                                 ifelse(test_condition == "unce", "uncertain",
                                        test_condition)))

d_test %>% 
  ggplot(., aes(x = test_condition, y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1))
# add mean / CI
```

Plot main rel / comp / pri questions by main condition

```{r}
d_critical <- d_critical %>% 
  pivot_longer(c(competence, relevance, prior), 
               names_to = "class_condition", 
               values_to = "prior_class")
d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  ggplot(., aes(x = as.factor(prior_class), y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
    facet_wrap(main_type~class_condition)
```
Plot main rel / comp questions by main condition, separated into with / without critical utterance
```{r}
# add info to already record this
d_critical <- d_critical %>% 
  mutate(w_utterance = ifelse(is.na(critical_question), F, T),
         block = ifelse(block == "comp", "competence", 
                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))

d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
    facet_wrap(main_type~class_condition) # get ratings from the respective trials only 
# use block for getting the actual correct ratings, but class_condition for subsetting the prior classification fo the respective predictor

# # some condition was wrongly recorded as pri
# d_critical %>% 
#      filter(block != "xor" & block != "some") %>%
#      filter(block == class_condition) %>% filter(block == "prior" & w_utterance == T) %>%
#   View()

d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
  summarize(mean_response = mean(response))
```

plot inference rating as a function of respective rating of the explanatory factor (think about a plot where one would see an interaction)