### 2021-03-28 analysis/xor-some-prolific-pilot1.Rmd
check by-participant performance

---
title: "XOR-Some Prolific Pilot 1"
author: "Polina Tsvilodub"
date: "3/24/2021"
output: github_document
---

In this first pilot on Prolific we test the implementation of the xor study in magpie. The structure of the experiment was as follows:
Participants read instructions, completed three example trials, and then completed 8 main blocks consisting of 4 xor and 4 some items. Each main block had the following structure: Participants read the background story, answered one comprehension question, then answered competence / relevance / prior questions in randomized order; then they read another 3 comprehension questions, after which the critical utterance was added below the background story. They answered the inference strength question, and then competence / relevance questions in randomized order again.

N=10 participants were recruited for this pilot and compensated 2 pounds/participant. 8 items (4 some + 4 xor) were manually selected for this pilot to get more data points per item.

```{r libraries}
library(tidyverse)
library(tidyboot)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)
```

```{r anonymization, include=FALSE}
#d_raw <- read_csv("~/Documents/Research/XOR/results_58_xor-some-Prolific-pilot1_N10.csv")
# remove IDs
#d_raw %>% select(-prolific_id) %>% write_csv("../data/pilots/results_58_xor-some-Prolific-pilot1_N10.csv") 
```

```{r data, results='hide'}
d <- read_csv("../data/pilots/results_58_xor-some-Prolific-pilot1_N10.csv")
```

Checking if there are any comments indicating technical issues:
```{r comments}
d %>% distinct(comments)
```

Check native languages. In the main experiment, participants not indicating English as (one of) their native language(s) will be excluded. We also compute some participant demographics.
```{r languages}
d %>% distinct(languages)
# exclude non-natives if necessary
d_native <- d

d_native %>% pull(age) %>% mean(., na.rm = T) 
d_native %>% count(gender) %>% mutate(n = n/88)
```

Next, we check whether all the conditions were used correctly.

```{r counts}
# check xor/some vs. trial type
d_native %>% count(main_type, condition) 
# check xor/some vs. experimental condition
d_native %>% count(main_type, exp_condition)
# count items used
d_native %>% count(title)
```

Check the time participants spent overall on the experiment:
```{r responseTime}
overall_timeSpent <- d_native %>% mutate(timeSpent = round(timeSpent, 2)) %>% distinct(timeSpent)

#  summarize(timeCounts = count(timeSpent) / d_native %>% )
ggplot(data = overall_timeSpent, aes(y=timeSpent, alpha = 0.7)) +
  geom_boxplot() +
  ggtitle("Overall time participants took")
```

Next, we would exclude participants based on their ratings in the main / example trials (and possibly the bot check) according to preregistered exclusion criteria. For now, all the data is considered.
```{r clean}
d_main <- d_native %>% select(-age, -botresponse, -comments, -education, -endTime, 
                              -gender, -languages, -optionLeft, -optionRight, -startDate,
                              -startTime, -timeSpent) %>%
  filter(trial_name != "example")
d_exmpl <- d_native %>% select(-age, -botresponse, -comments, -education, -endTime, 
                              -gender, -languages, -optionLeft, -optionRight, -startDate,
                              -startTime, -timeSpent) %>%
  filter(trial_name == "example")
d_critical <- d_main %>% filter(condition == "critical")

# get overall mean ratings / subject
d_native %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)
```
Plot responses on example questions by question type

Plot test questions by type
```{r}
d_test <- d_main %>% rowwise() %>% filter(condition == "test") %>% 
  mutate(test_condition = substr(test_question, 6, 9),
         test_condition = ifelse(test_condition == "fals", "false", 
                                 ifelse(test_condition == "unce", "uncertain",
                                        test_condition)))
d_test_ci <- d_test %>% group_by(test_condition) %>% 
  tidyboot_mean(column = response)

d_test %>% 
  ggplot(., aes(x = test_condition, y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
  geom_point(data = d_test_ci, aes(x = test_condition, y = mean), color = "red", 
             size = 4) 
```

```{r, include=FALSE}
# Plot main rel / comp / pri questions by main condition. 
d_critical <- d_critical %>% 
  pivot_longer(c(competence, relevance, prior), 
               names_to = "class_condition", 
               values_to = "prior_class")
d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  ggplot(., aes(x = as.factor(prior_class), y = response, color = as.factor(prior_class) )) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
    facet_wrap(main_type~class_condition) +
  xlab("Anticipated categorization (high / low)") +
  theme(legend.position = 'none') 
```
Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis), separated into with / without critical utterance (shape, color). Only high-competence items were used for the some trials (only low-competence in the xor trials, respectively). This will be balanced in the main study. 
```{r}
# add info to already record this
d_critical <- d_critical %>% 
  mutate(w_utterance = ifelse(is.na(critical_question), F, T),
         block = ifelse(block == "comp", "competence", 
                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
  summarize(mean_response = mean(response)) -> d_critical_summary

d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response, shape = w_utterance, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
             color = "red", size = 3) +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) # get ratings from the respective trials only

```

Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
``` {r}
d_critical %>% 
  filter(block == "xor" | block == "some") %>%
  #group_by(block, class_condition, prior_class) %>%
  #summarize(mean_rating = mean(response)) %>%
  ggplot(., aes(x = prior_class, y = response)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_line() +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Mean inference strength ratings by-predictor")

```

Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given.

``` {r}
# extending 'conditions' labels to include wheter the utterance was present or not
d_critical <- d_critical %>% 
  mutate(block_extended = ifelse(
    !w_utterance, 
    block, 
    ifelse(block %in% c("some", "xor"), "target", str_c(block, "_wUtt", ""))
  ))
```

Plot mean ratings (across all 10 participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition. For relevance and competence, the color indicates whether it was presented without the utterance or with. 
```{r, fig.width=10, fig.height=7}
bar.width = 0.8
d_critical %>% 
  mutate(title = paste(title, exp_condition, sep = "_")) %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  group_by(block_extended, title, w_utterance, main_type) %>%
  summarize(mean_rating = mean(response)) %>% 
  ggplot(., aes(x = block_extended, y = mean_rating, fill = w_utterance)) +
  geom_col(alpha = 0.7, width = bar.width, position = position_dodge(width = bar.width)) +
  #geom_point(size = 2, alpha = 0.5, position = position_jitter(width = 0.1)) +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question mean ratings")
```

``` {r}
# make a versatile wide representation of the critical data
d_critical_wide <- d_critical %>% 
  select(submission_id, title, main_type, block_extended, response) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )

# correlation plot for "some"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "some") %>%  
    select(prior:ncol(d_critical_wide))
)

# correlation plot for "or"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "xor") %>%  
    select(prior:ncol(d_critical_wide))
  )
```

Check by-participant behavior to see if they show versatile response strategies for the different questions for xor vs. some. Participants show a sufficient range of responses across conditions.  
```{r, fig.width=15, fig.height=7}
d_critical %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  ggplot(., aes(x = block, y = response, color = block)) +
  geom_point(alpha = 0.7, size = 2) + 
  facet_wrap(main_type~submission_id, ncol = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-participant ratings all xor vs. some trials")
```

In conclusion, the technical side of the experiment seems to work without glitches; participants perform as expected on the comprehension questions; the relevance / competence / prior ratings seem to be mostly in line with our anticipated categorizations.

One point to note is that on average participants took slightly longer than the expected 15 minutes to complete the experiment.