### 2021-04-04 analysis/xor-some-prolific-pilot2.Rmd
add stats

---
title: "XOR-Some Prolific Pilot 2"
author: "Polina Tsvilodub"
date: "4/2/2021"
output: github_document
---

In this second pilot on Prolific we gather a more substantial amount of pilot data for the xor-some study on Prolific. The structure of the experiment was as follows:
Participants read instructions, completed three example trials, and then completed 8 main blocks consisting of 4 xor and 4 some items. Each main block had the following structure: Participants read the background story, answered one comprehension question, then answered competence / relevance / prior questions in randomized order; then they read another 3 comprehension questions, after which the critical utterance was added below the background story. They answered the inference strength question, and then competence / relevance questions in randomized order again.

N=118 participants were recruited for this pilot and compensated 2 pounds/participant. 8 items (4 some + 4 xor) were sampled at random for each participant such that each participant saw one item in each condition (relevance X competence X prior = 8 unique conditions).

```{r libraries, include=FALSE}
library(tidyverse)
library(tidyboot)
library(brms)
library(tidybayes)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)
```

```{r anonymization, include=FALSE}
#d_raw <- read_csv("~/Documents/Research/XOR/raw_data/results_58_xor-some-Prolific-pilot2_N120.csv")
# remove IDs
#d_raw %>% select(-prolific_id) %>% write_csv("../data/pilots/results_58_xor-some-Prolific-pilot2_N120.csv") 
```

## Checks & Exclusions
```{r data, results='hide', include=FALSE}
d <- read_csv("../data/pilots/results_58_xor-some-Prolific-pilot2_N120.csv")
```

Checking if there are any comments indicating technical issues which is not the case:
```{r comments}
d %>% distinct(comments) %>% View()
```

Check native languages. Participants not indicating English as (one of) their native language(s) are be excluded. We also compute some participant demographics.
```{r languages}
d %>% distinct(languages)
cat("Number of partiipants before excluding non-natives: ", d %>% distinct(submission_id) %>% count() %>% pull() ) 
# exclude non-natives if necessary
d_native <- d %>% 
 # filter(("en" | "En" | "bri") %in% languages)
  filter(grepl("[(en)(br)]", languages, ignore.case = T))

cat("Number of partiipants after excluding non-natives: ", d_native %>% distinct(submission_id) %>% count() %>% pull() )

cat("Mean age: ", d_native %>% pull(age) %>% mean(., na.rm = T) )
d_native %>% count(gender) %>% mutate(n = n/88)
```

Next, we check whether all the conditions were used correctly.

```{r counts}
# check xor/some vs. trial type
d_native %>% count(main_type, condition) 
# check xor/some vs. experimental condition
d_native %>% count(main_type, exp_condition)
# count items used
d_native %>% count(title)
```

Check the time participants spent overall on the experiment before cleaning the data:
```{r responseTime}
overall_timeSpent <- d_native %>% mutate(timeSpent = round(timeSpent, 2)) %>% distinct(timeSpent)

#  summarize(timeCounts = count(timeSpent) / d_native %>% )
ggplot(data = overall_timeSpent, aes(y=timeSpent, alpha = 0.7)) +
  geom_boxplot() +
  ggtitle("Overall time participants took in mins")
```
```{r}
d_main <- d_native %>% select(-age, -botresponse, -comments, -education, -endTime, 
                              -gender, -languages, -optionLeft, -optionRight, -startDate,
                              -startTime, -timeSpent) %>%
  filter(trial_name != "example")
d_exmpl <- d_native %>% select(-age, -botresponse, -comments, -education, -endTime, 
                              -gender, -languages, -optionLeft, -optionRight, -startDate,
                              -startTime, -timeSpent) %>%
  filter(trial_name == "example")
d_critical <- d_main %>% filter(condition == "critical")
```

Plot responses on comprehension questions by type before applying exclusion criteria:
```{r, echo=FALSE}
d_test <- d_main %>% rowwise() %>% filter(condition == "test") %>% 
  mutate(test_condition = substr(test_question, 6, 9),
         test_condition = ifelse(test_condition == "fals", "false", 
                                 ifelse(test_condition == "unce", "uncertain",
                                        test_condition)))
d_test_ci <- d_test %>% group_by(test_condition) %>% 
  tidyboot_mean(column = response)

d_test %>% 
  ggplot(., aes(x = test_condition, y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
  geom_point(data = d_test_ci, aes(x = test_condition, y = mean), color = "red", 
             size = 4) +
  facet_wrap(~main_type)
```

Next, we exclude participants based on their ratings in the main trials: Participants who gave all responses within the range of 10 and participants who failed more than 0.2 of the comprehension questions are excluded from analysis. 
Participants who failed all example trials are excluded, as well. 
The bot check trial is not considered for exclusions. 
```{r exclusions}
# get participants failing example trials
d_exmpl <- d_exmpl %>% group_by(submission_id) %>% 
  mutate(example_condition = ifelse(grepl("as certainly true", question), "true", 
                                    ifelse(grepl("as certainly false", question), "false",
                                           "uncertain")),
         passed_example_trial = case_when(example_condition == "true" ~ response >= 80,
                                    example_condition == "false" ~ response <= 20,
                                    example_condition == "uncertain" ~ response >= 30),
       #  check if all trials passed
         passed_example = case_when(sum(passed_example_trial) == 0 ~ FALSE,
                                    TRUE ~ TRUE)
         ) %>% filter(passed_example == F)

cat("Subjects failing the example trials: ", d_exmpl %>% distinct(submission_id) %>% pull() %>% length())

# apply exclusion criteria to main trials
# check range of responses per participant
d_main_fail <- d_main %>% group_by(submission_id) %>%
  mutate(passed_main = case_when(max(response) - min(response) <= 10 ~ FALSE,
                                 TRUE ~ TRUE)
         ) %>% filter(passed_main == F)
cat("Subjects providing the same ratings throughout the trials: ", d_main_fail %>% distinct(submission_id) %>% pull() %>% length())

# get participants failing comprehension questions
d_test <- d_test %>%
  group_by(submission_id) %>%
  mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 70,
                                   test_condition == "false" ~ response <= 40,
                                   test_condition == "uncertain" ~ response %in% (0:90)),
         mean_comprehension = mean(passed_filler_trial),
         passed_filler = mean_comprehension >= 0.8
         ) %>%
  filter(passed_filler == F)

cat("Subjects failing the comprehension trials: ", d_test %>% distinct(submission_id) %>% pull() %>% length())

# put it all together
d_full_clean <- anti_join(d_main, d_main_fail, by = "submission_id")
d_full_clean <- anti_join(d_full_clean, d_exmpl, d_test, by = "submission_id")
d_full_clean <- anti_join(d_full_clean, d_test, by = "submission_id")

cat("Nr. of participants left after cleaning: ", d_full_clean %>% distinct(submission_id) %>% pull() %>% length())
```

```{r clean}
# get overall mean ratings / subject
d_full_clean %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)
```

```{r, echo=FALSE}
d_critical_clean <- d_full_clean %>% filter(condition == "critical")
d_critical_long <- d_critical_clean %>% 
  pivot_longer(c(competence, relevance, prior), 
               names_to = "class_condition", 
               values_to = "prior_class")
```

## Plots
Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis), separated into with / without critical utterance (shape, color). The prior questions were only used without the utterance. 
```{r}

d_critical_long <- d_critical_long %>% 
  mutate(w_utterance = ifelse(is.na(critical_question), F, T),
         block = ifelse(block == "comp", "competence", 
                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
  summarize(mean_response = mean(response)) -> d_critical_summary

d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response, shape = w_utterance, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
             color = "red", size = 3) +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) # get ratings from the respective trials only

```
```{r}
# check whether the ratings in the two prior questions are the same in xor 
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  filter(block == "prior", main_type == "xor") %>%
  mutate(priorQ_nr = rep(c(1,2), 396)) -> d_xor_priors
d_xor_priors %>% group_by(priorQ_nr, prior_class) %>%
  summarise(mean = mean(response)) -> d_xor_priors_summary
d_xor_priors %>%
  ggplot(., aes(x = as.factor(priorQ_nr), y = response )) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_xor_priors_summary, aes(x = as.factor(priorQ_nr), y = mean), color = "red", size = 3) +
  xlab("First vs Second prior question for high vs low prior conditions") +
  facet_wrap(~as.factor(prior_class)) # get ratings from the respective trials only

```

Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
``` {r}
d_critical_long %>% 
  filter(block == "xor" | block == "some") %>%
  #group_by(block, class_condition, prior_class) %>%
  #summarize(mean_rating = mean(response)) %>%
  ggplot(., aes(x = prior_class, y = response)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_smooth(method="lm") +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Inference strength ratings by-predictor")

```
Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given.

``` {r}
# extending 'conditions' labels to include whether the utterance was present or not
d_critical_long <- d_critical_long %>% 
  mutate(block_extended = ifelse(
    !w_utterance, 
    block, 
    ifelse(block %in% c("some", "xor"), "target", str_c(block, "_wUtt", ""))
  ))
```

```{r}
# make a versatile wide representation of the critical data
d_critical_wide <- d_critical_long %>% 
  select(submission_id, title, main_type, block_extended, response) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response, 
    values_fn = mean # getting means for double prior measurement in "xor"
  ) 
```

Plot inference strength ratings against raw predictor ratings for each participant / each item:
```{r}
d_critical_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

d_critical_wide %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp

d_critical_wide %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri

gridExtra::grid.arrange(p.rel, p.comp, p.pri, ncol = 3) 
```

Plot mean ratings (across all participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition. For relevance and competence, the color indicates whether it was presented without the utterance or with. 
```{r, fig.width=10, fig.height=30}
bar.width = 0.8
d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_")) %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  group_by(block_extended, title, w_utterance, main_type) %>%
  summarize(mean_rating = mean(response)) %>% 
  ggplot(., aes(x = block_extended, y = mean_rating, fill = w_utterance)) +
  geom_col(alpha = 0.7, width = bar.width, position = position_dodge(width = bar.width)) +
  #geom_point(size = 2, alpha = 0.5, position = position_jitter(width = 0.1)) +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question mean ratings")
```

``` {r}

# correlation plot for "some"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "some") %>%  
    select(prior:ncol(d_critical_wide))
)

# correlation plot for "or"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "xor") %>%  
    select(prior:ncol(d_critical_wide))
  )
```

Check by-participant behavior to see if they show versatile response strategies for the different questions for xor vs. some. Participants show a sufficient range of responses across conditions.  
```{r, fig.width=15, fig.height=7}
d_critical_long %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  ggplot(., aes(x = block, y = response, color = block)) +
  geom_point(alpha = 0.7, size = 2) + 
  facet_wrap(main_type~submission_id, ncol = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-participant ratings all xor vs. some trials")
```

## Explore z-scoring
Visual exploration of z-scoring each measure (rel / comp / pri / target / comp-wUtterance / rel-wUtterance) per participant, collapsing some and or together. 
```{r}
d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd)

d_critical_zScored %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
  summarize(mean_response = mean(response_centered)) -> d_critical_zScore_summary

d_critical_zScored %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response_centered, shape = w_utterance, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_zScore_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
             color = "red", size = 3) +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) + # get ratings from the respective trials only 
  ggtitle("Centered responses to by-predictor questions")
```

Plot z-scored predictor ratings against z-scored target ratings
```{r}
d_critical_zScored_wide <- d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )

d_critical_zScored_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  facet_wrap(~main_type, ncol = 1) -> p.rel.z

d_critical_zScored_wide %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp.z

d_critical_zScored_wide %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri.z

gridExtra::grid.arrange(p.rel.z, p.comp.z, p.pri.z, ncol = 3) 
```

## Stats
Maximal models on raw ratings:
```{r, results='hide'}
# xor, maximal model with interactions and maximal REs
model_xor <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor)
```

```{r, results='hide'}
# some, maximal model with interactions and maximal REs
model_some <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some)
```

#### Stats on z-scored data
```{r, results='hide'}
# xor, maximal model with interactions and maximal REs on z-scored data
model_xor_zScored <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored)
```

```{r, results='hide'}
# some, maximal model with interactions and maximal REs
model_some_zScored <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored)
```

#### Stats with dichotomized predictors 
Maximal model on raw inference ratings, using anticipated binary categorization of items as predictors 
```{r, results='hide'}
d_critical_long %>% 
  select(submission_id, title, main_type, block_extended, response, prior_class, class_condition) %>% 
  unique() %>% 
  pivot_wider(
    names_from = class_condition, 
    values_from = prior_class
  )  %>%
  mutate(relevance = as.factor(relevance),
         competence = as.factor(competence),
         prior = as.factor(prior)) %>% 
  filter(block_extended == "target") -> d_critical_wide_cat
# sum code predictors
# low prior: 1, high prior:: -1 
contrasts(d_critical_wide_cat$prior) <- contr.sum(2)
# low comp : 1, high comp : -1
contrasts(d_critical_wide_cat$competence) <- contr.sum(2)
# low rel: 1, high rel: -1
contrasts(d_critical_wide_cat$relevance) <- contr.sum(2)

# xor, maximal model with interactions and maximal REs
model_xor_cat <- brm(
  response ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide_cat %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_cat)
```

```{r, results='hide'}
model_some_cat <- brm(
  response ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide_cat %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_some_cat)
```

Maximal model on z-scored inference ratings, using anticipated binary categorization of items as predictors 
```{r, results='hide'}
d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered, prior_class, class_condition) %>% 
  unique() %>% 
  pivot_wider(
    names_from = class_condition, 
    values_from = prior_class
  )  %>%
  mutate(relevance = as.factor(relevance),
         competence = as.factor(competence),
         prior = as.factor(prior)) %>% 
  filter(block_extended == "target") -> d_critical_wide_cat_zScored
# sum code predictors
# low prior: 1, high prior:: -1 
contrasts(d_critical_wide_cat_zScored$prior) <- contr.sum(2)
# low comp : 1, high comp : -1
contrasts(d_critical_wide_cat_zScored$competence) <- contr.sum(2)
# low rel: 1, high rel: -1
contrasts(d_critical_wide_cat_zScored$relevance) <- contr.sum(2)

# xor, maximal model with interactions and maximal REs
model_xor_cat_zScored <- brm(
  response_centered ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_cat_zScored)
```

```{r, results='hide'}
model_some_cat_zScored <- brm(
  response_centered ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_some_cat_zScored)
```

In summary, the different models provide somewhat different results, specially when using raw ratings vs. binary predictors. Since interpreting the categorical models in relation to the plots might be not as straight-forward, I am not entirely sure whether the qualitative results match in all models. Using centered data seems to provide a little tighter estimates in the models; not sure where the slight qualitative differences come from.