### 2021-07-25 analysis/xor-some-prolific-pilot3.Rmd
analyse pilot 3 with updated vignettes

---
title: "XOR-Some Prolific Pilot 3"
author: "Polina Tsvilodub"
date: "7/24/2021"
output: github_document
---

In this third pilot on Prolific we gather pilot data for the vignettes that were substantially updated based on the results from the [second pilot](https://github.com/magpie-ea/magpie-xor-experiment/blob/master/analysis/xor-some-prolific-pilot2.md). The used vignettes were selected based on manual comparison and comprise 38 out of 64 vignettes.

The structure of the experiment was as follows:
Participants read instructions, completed three example trials, and then completed 8 main blocks consisting of 4 xor and 4 some items. Each main block had the following structure: Participants read the background story, answered one comprehension question, then answered competence / relevance / prior questions in randomized order; then they read another 3 comprehension questions, after which the critical utterance was added below the background story. They answered the inference strength question. The second round of relevance/competence questions was removed for the pilot.

N=50 participants were recruited for this pilot and compensated 2 pounds/participant. 8 items (4 some + 4 xor) were sampled at random for each participant such that each participant saw one item in each condition (relevance X competence X prior = 8 unique conditions).

```{r libraries, include=FALSE}
library(tidyverse)
library(tidyboot)
library(brms)
library(tidybayes)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)
```

```{r anonymization, include=FALSE}
#d_raw <- read_csv("~/Documents/Research/XOR/magpie-xor-experiment/data/pilots/results_63_xor-some-Prolific-pilot3_N50.csv")
# remove IDs
# d_raw %>% select(-prolific_id) %>% write_csv("../data/pilots/results_63_xor-some-Prolific-pilot3_N50.csv") 
```

## Checks & Exclusions
```{r data, results='hide', include=FALSE}
d <- read_csv("../data/pilots/results_63_xor-some-Prolific-pilot3_N50.csv")
```

Checking if there are any comments indicating technical issues which is not the case:
```{r comments}
d %>% distinct(comments) %>% View()
```

Check native languages. Participants not indicating English as (one of) their native language(s) are be excluded. We also compute some participant demographics.
```{r languages}
d %>% distinct(languages)
cat("Number of participants before excluding non-natives: ", d %>% distinct(submission_id) %>% count() %>% pull() ) 
# exclude non-natives if necessary
d_native <- d %>% 
 # filter(("en" | "En" | "bri") %in% languages)
  filter(grepl("[(en)(br)]", languages, ignore.case = T))

cat(" Number of participants after excluding non-natives: ", d_native %>% distinct(submission_id) %>% count() %>% pull() )

cat(" Mean age: ", d_native %>% pull(age) %>% mean(., na.rm = T) )
d_native  %>% count(gender) %>% mutate(n = n/72)
```

Next, we check whether all the conditions were used correctly.

```{r counts}
# check xor/some vs. trial type; there are more xor critical conditions because there are 2 prior questions
d_native %>% count(main_type, condition) 
# check experimental conditions
d_native %>% count(exp_condition)
# check xor/some vs. experimental condition
d_native %>% count(main_type, exp_condition)
# count items used
d_native %>% count(title)
```

Check the time participants spent overall on the experiment before cleaning the data:
```{r responseTime}
overall_timeSpent <- d_native %>% mutate(timeSpent = round(timeSpent, 2)) %>% distinct(timeSpent)

#  summarize(timeCounts = count(timeSpent) / d_native %>% )
ggplot(data = overall_timeSpent, aes(y=timeSpent, alpha = 0.7)) +
  geom_boxplot() +
  ggtitle("Overall time participants took in mins")
```

Investigate very fast completion times (we have none in this pilot):
```{r}
 d_native %>% mutate(timeSpent = round(timeSpent, 2)) %>%
  filter(timeSpent < 8) -> fast_completions

# check if their responses look somehow peculiar
#fast_completions %>%
#  ggplot(., aes(x = response)) +
#  geom_histogram()

cat("number of fast-responders (below 8 mins):", fast_completions %>% distinct(submission_id) %>% pull() %>% length())
```
Data frames preparation:
```{r}
d_main <- d_native %>% select(-age, -botresponse, -comments, -education, -endTime, 
                              -gender, -languages, -optionLeft, -optionRight, -startDate,
                              -startTime, -timeSpent) %>%
  filter(trial_name != "example")
d_exmpl <- d_native %>% select(-age, -botresponse, -comments, -education, -endTime, 
                              -gender, -languages, -optionLeft, -optionRight, -startDate,
                              -startTime, -timeSpent) %>%
  filter(trial_name == "example")
d_critical <- d_main %>% filter(condition == "critical")
```

Plot responses on comprehension questions by type before applying exclusion criteria:
```{r, echo=FALSE}
d_test <- d_main %>% rowwise() %>% filter(condition == "test") %>% 
  mutate(test_condition = substr(test_question, 6, 9),
         test_condition = ifelse(test_condition == "fals", "false", 
                                 ifelse(test_condition == "unce", "uncertain",
                                        test_condition)))
d_test_ci <- d_test %>% group_by(test_condition) %>% 
  tidyboot_mean(column = response)

d_test %>% 
  ggplot(., aes(x = test_condition, y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
  geom_point(data = d_test_ci, aes(x = test_condition, y = mean), color = "red", 
             size = 4) +
  facet_wrap(~main_type)
```

Next, we exclude participants based on their ratings in the main trials: Participants who gave all responses within the range of 10 and participants who failed more than 0.2 of the comprehension questions are excluded from analysis. 
Participants who failed all example trials are excluded, as well. 
The bot check trial is not considered for exclusions. 
```{r exclusions}
# get participants failing example trials
d_exmpl_fail <- d_exmpl %>% group_by(submission_id) %>% 
  mutate(example_condition = ifelse(grepl("as certainly true", question), "true", 
                                    ifelse(grepl("as certainly false", question), "false",
                                           "uncertain")),
         passed_example_trial = case_when(example_condition == "true" ~ response >= 80,
                                    example_condition == "false" ~ response <= 20,
                                    example_condition == "uncertain" ~ response >= 30),
       #  check if all trials passed
         passed_example = case_when(sum(passed_example_trial) == 0 ~ FALSE,
                                    TRUE ~ TRUE)
         ) %>% filter(passed_example == F)

cat("Subjects failing the example trials: ", d_exmpl_fail %>% distinct(submission_id) %>% pull() %>% length())

# apply exclusion criteria to main trials
# check range of responses per participant
d_main_fail <- d_main %>% group_by(submission_id) %>%
  mutate(passed_main = case_when(max(response) - min(response) <= 10 ~ FALSE,
                                 TRUE ~ TRUE)
         ) %>% filter(passed_main == F)
cat(" Subjects providing the same ratings throughout the trials: ", d_main_fail %>% distinct(submission_id) %>% pull() %>% length())

# get participants failing comprehension questions
d_test <- d_test %>%
  group_by(submission_id) %>%
  mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 60,
                                   test_condition == "false" ~ response <= 40,
                                   test_condition == "uncertain" ~ response %in% (0:90)),
         mean_comprehension = mean(passed_filler_trial),
         passed_filler = mean_comprehension >= 0.8
         )

d_test_fail <- d_test %>% 
  filter(passed_filler == F)

cat(" Subjects failing the comprehension trials: ", d_test_fail %>% distinct(submission_id) %>% pull() %>% length())

# plot by-participant means on comprehension questions, and mark red excluded points, 
# to see if the participants performed in a peculiar way
d_test %>% 
  group_by(submission_id, main_type, test_condition) %>%
  mutate(subj_mean = mean(response)) %>%
  ggplot(., aes(x = test_condition, y = subj_mean)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
  geom_point(data = . %>% filter(passed_filler == F), aes(x = test_condition, y = subj_mean), color = "red", 
             size = 2) +
  facet_wrap(~main_type) +
  ggtitle("Red points mark subject response means for subjects\nwith a mean comprehension question accuracy < 0.8")
```

```{r}
# put it all together
d_full_clean <- anti_join(d_main, d_main_fail, by = "submission_id")
d_full_clean <- anti_join(d_full_clean, d_exmpl_fail, d_test, by = "submission_id")
d_full_clean <- anti_join(d_full_clean, d_test_fail, by = "submission_id")

cat(" Nr. of participants left after cleaning: ", d_full_clean %>% distinct(submission_id) %>% pull() %>% length())
```

```{r clean}
# get overall mean ratings / subject across comprehension and critical trials 
d_full_clean %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)

# get mean ratings / subject in critical trials 
d_full_clean %>% filter(condition == "critical") %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)

```

```{r, echo=FALSE}
d_critical_clean <- d_full_clean %>% filter(condition == "critical")
d_critical_long <- d_critical_clean %>% 
  pivot_longer(c(competence, relevance, prior), 
               names_to = "class_condition", 
               values_to = "prior_class")

d_critical_long <- d_critical_long %>% 
  mutate(w_utterance = ifelse(is.na(critical_question), F, T),
         block = ifelse(block == "comp", "competence", 
                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
```

Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given (no such condition in this pilot).

``` {r}
# extending 'conditions' labels to include whether the utterance was present or not
d_critical_long <- d_critical_long %>% 
  mutate(block_extended = ifelse(
    !w_utterance, 
    block, 
    ifelse(block %in% c("some", "xor"), "target", str_c(block, "_wUtt", ""))
  ))
```

Average nr of responses per question per vignette:
```{r}
d_critical_long %>% 
  filter(class_condition == block | block == "xor" | block == "some") %>%
  select(-class_condition, -prior_class) %>%
  unique() %>% count(title, block_extended, main_type) %>%
  mutate(n = ifelse(main_type == "xor" & block_extended == "prior", n/2, n)) %>%
  summarize(mean_observations = mean(n))
```

Compute the average deviation from expected responses for each vignette, individually for each dimension (rel / comp / pri). Only half as many items deviate from expectations by >= 50 points compared to the last pilot.
```{r}
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  mutate(expected_rating = prior_class * 100,
         response_deviation = abs(response - expected_rating),
         block = ifelse(block == "xor" | block == "some", "target", block)
         ) %>%
  group_by(title, block) %>%
  summarize(mean_deviation = mean(response_deviation)) %>%
  arrange(desc(mean_deviation), .by_group = T) %>%
  # get the largest deviations - over 50 
  filter(mean_deviation >= 50) -> d_critical_deviations

# d_critical_deviations %>% write_csv("../data/pilots/pilot3_byItem_rating_deviations.csv")

print(d_critical_deviations)

# count which dimension is the most problematic - the prior
d_critical_deviations %>% group_by(block) %>% count()

# check which items have the most deviations over 50 
d_critical_deviations %>% group_by(title) %>% count() %>% arrange(desc(.$n))
```

Compute the standard deviation for each question for each item as an additional measure of spread in the participants' responses within-item:
```{r}
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  mutate(
         block = ifelse(block == "xor" | block == "some", "target", block)
         ) %>%
  group_by(title, block) %>%
  summarize(item_sd = sd(response)) %>%
  arrange(desc(item_sd), .by_group = T) %>%
  # arbitrary set to 25
  filter(item_sd >= 25) -> d_critical_sd
d_critical_sd

# d_critical_sd %>% write_csv("../data/pilots/pilot3_byItem_rating_SDs.csv")
```

Get the vignettes and dimensions that were strongly deviating from expectations in the previous pilot and still did not improve:
```{r}
deviating_item_pilot2 <- read_csv("../data/pilots/pilot2_byItem_rating_deviations.csv") %>% select(-mean_deviation)
d_critical_deviations %>% select(-mean_deviation) %>% semi_join(., deviating_item_pilot2)
```

Visually check the spread of by-item ratings among subjects for each question:
```{r, echo=FALSE, fig.height=25, fig.width=10}
d_critical_long_grandMeans <- d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_"),
         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
  filter(block == class_condition | (block == "target" & class_condition == "relevance")) %>%
  group_by(main_type, block) %>% summarize(mean_resp = mean(response))

d_critical_long_expectedCats <- d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_"),
         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
  filter(block == class_condition | (block == "target" & class_condition == "relevance")) %>%
  group_by(main_type, block, title) %>% summarize(mean_resp_block = mean(response))

d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_"),
         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
  left_join(., d_critical_long_grandMeans, by = c("main_type", "block")) %>% 
  filter(block == class_condition | ((block == "target") & (class_condition == "relevance"))) -> d_long_plot_data

left_join(d_long_plot_data, d_critical_long_expectedCats, by = c("main_type", "block", "title")) %>% 
  mutate(expected_block = ifelse(block != "target", prior_class * 100, NA),
         block = factor(block, levels = c("relevance", "competence", "prior", "target"))) -> d_long_plot_full
d_long_plot_full %>%
  ggplot(., aes(x = block, y = response, color = block)) +
  geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
  geom_hline(aes(yintercept = mean_resp), alpha = 0.7) +
  geom_point(aes(y = mean_resp), color = "black", shape = 2) +
  geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
  geom_point(aes(y = expected_block), color="red", size = 2.5, na.rm = T) +
  ylab("Raw responses to respective questions") +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question raw ratings\nBlack dots indicate by-block means within-item, lines and triangles indicate grand by-block means\nRed dots indicate expected rating")
```

Visually compare the updated vignettes (pilot 2 vs current pilot)
```{r, echo=FALSE}
# add column indicating which pilot it is
d_long_plot_full <- d_long_plot_full %>%
  mutate(pilot = "updated")

d_long_plot_pilot2 <- read_csv("../data/pilots/pilot2_byItem_plotting.csv") %>% 
  mutate(pilot = "old")
# only keep data for those items that were also tested in the current pilot
d_long_plot_pilot2 <- semi_join(d_long_plot_pilot2, d_long_plot_full, by = c("title"))

d_long_plot_both_pilots <- bind_rows(d_long_plot_full, d_long_plot_pilot2)
```

```{r, echo=FALSE, fig.width=8, fig.height=15}
# plot old and updated some items side-by-side
# the vignettes for which only an "updated" plot occurs are new items that were added instead of some other vignettes 
d_long_plot_both_pilots %>%
  filter(main_type == "some") %>%
  ggplot(., aes(x = block, y = response, color = block)) +
  geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
  geom_hline(aes(yintercept = mean_resp), alpha = 0.7) +
  geom_point(aes(y = mean_resp), color = "black", shape = 2) +
  geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
  geom_point(aes(y = expected_block), color="red", size = 2.5, na.rm = T) +
  ylab("Raw responses to respective questions") +
  facet_wrap(title~pilot, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question raw ratings, old vs. updated SOME vignettes\nBlack dots indicate by-block means within-item, lines and triangles indicate grand by-block means\nRed dots indicate expected rating")
```

```{r, echo=FALSE, fig.width=8, fig.height=15}
# plot old and updated xor items side-by-side 
d_long_plot_both_pilots %>%
  filter(main_type == "xor") %>%
  ggplot(., aes(x = block, y = response, color = block)) +
  geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
  geom_hline(aes(yintercept = mean_resp), alpha = 0.7) +
  geom_point(aes(y = mean_resp), color = "black", shape = 2) +
  geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
  geom_point(aes(y = expected_block), color="red", size = 2.5, na.rm = T) +
  ylab("Raw responses to respective questions") +
  facet_wrap(title~pilot, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question raw ratings, old vs. updated XOR vignettes\nBlack dots indicate by-block means within-item, lines and triangles indicate grand by-block means\nRed dots indicate expected rating")
```

Check by-vignette by-predictor empirical means, to see if the range is somewhat covered:
```{r}
d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_"),
         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
  filter(block == class_condition | block == "target") %>%
  group_by(title, block) %>%
  summarise(mean_response = mean(response)) %>%
  arrange(mean_response)
```

## Plots
```{r, echo=FALSE, eval=FALSE}
# Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis), separated into with / without critical utterance (shape, color). The prior questions were only used without the utterance. 
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
  summarize(mean_response = mean(response)) -> d_critical_summary

d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_summary, aes(x = as.factor(prior_class), y = mean_response), 
             color = "red", size = 3) +
  ylab("Responses to respective predictor question") +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) # get ratings from the respective trials only

```
First, compare the two prior ratings in the xor condition:
```{r, fig.height=3}
# check whether the ratings in the two prior questions are the same in xor 
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  filter(block == "prior", main_type == "xor") %>%
  mutate(priorQ_nr = rep(c(1,2), nrow(.)/2),
         unique_ID = paste(submission_id, ID, sep="_")) -> d_xor_priors
d_xor_priors %>% group_by(priorQ_nr, prior_class) %>%
  summarise(mean = mean(response)) -> d_xor_priors_summary

d_xor_priors %>%
  ggplot(., aes(x = as.factor(priorQ_nr), y = response )) +
  geom_point(size = 2, alpha = 0.6) +
  geom_point(data = d_xor_priors_summary, aes(x = as.factor(priorQ_nr), y = mean), color = "red", size = 3) +
  geom_path(aes(group = "unique_ID"), alpha = 0.6) +
  ylab("Responses to prior questions") +
  xlab("First vs Second prior question for high vs low prior conditions") +
  facet_wrap(~as.factor(prior_class)) + # get ratings from the respective trials only 
  ggtitle("Xor prior question ratings.\nLines indicate the two responses provided by one participant for a specific item.")
```

```{r}
# check analytically how many subjects rate prior questions differently
d_xor_priors %>% select(-trial_number, -RT, -prompt) %>% 
  pivot_wider(names_from = "priorQ_nr", values_from = "response") %>%
  mutate(response_difference = abs(`1` - `2`),
         bigDifference = ifelse(response_difference > 15, T, F)) -> d_xor_priors_differences

# 19 responses with difference > 15 between first and second rating
d_xor_priors_differences %>% count(bigDifference)

# proportion of big differences in prior responses by prior type (high vs low)
d_xor_priors_differences %>% group_by(prior_class) %>% summarise(prop_big_diffs = mean(bigDifference))
```

```{r, echo=FALSE, eval=FALSE}
# make a versatile wide representation of the critical data
d_critical_wide <- d_critical_long %>% 
  select(submission_id, title, main_type, block_extended, response) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response, 
    values_fn = mean # getting means for double prior measurement in "xor"
  ) 
```
```{r, echo=FALSE, eval=FALSE}
# Plot inference strength ratings against raw predictor ratings from each participant across items:

d_critical_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("Inference strength ratings") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

d_critical_wide %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp

d_critical_wide %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri

gridExtra::grid.arrange(p.rel, p.comp, p.pri, ncol = 3) 
```
```{r, fig.width=10, fig.height=30, echo=FALSE, eval=FALSE}
# Plot mean ratings (across all participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition.
bar.width = 0.8
d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_")) %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  group_by(block_extended, title, w_utterance, main_type) %>%
  summarize(mean_rating = mean(response)) %>% 
  ggplot(., aes(x = block_extended, y = mean_rating)) +
  geom_col(alpha = 0.7, width = bar.width, position = position_dodge(width = bar.width)) +
  #geom_point(size = 2, alpha = 0.5, position = position_jitter(width = 0.1)) +
  ylab("Mean responses to respective questions") +
  xlab("Question type") +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question mean ratings")
```
``` {r, echo=FALSE, eval=FALSE}

# correlation plot for "some"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "some") %>%  
    select(prior, competence, relevance, target) 
)

# correlation plot for "or"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "xor") %>%  
    select(prior, competence, relevance, target)
  )
```

As planned, all the primary visualizations and analyses below are conducted on data z-scored by-subject by-block.

Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis): 
```{r}
d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd,
         # catch the cases where sd is 0 
         response_centered = ifelse(is.na(response_centered), 0, response_centered))

d_critical_zScored %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
  summarize(mean_response = mean(response_centered)) -> d_critical_zScore_summary

d_critical_zScored %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response_centered, shape = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_zScore_summary, aes(x = as.factor(prior_class), y = mean_response), 
             color = "red", size = 3) +
  ylab("Responses to respective predictor questions") +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) + # get ratings from the respective trials only 
  ggtitle("Centered responses to by-predictor questions")
```
Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
``` {r}
d_critical_zScored %>% 
  filter(block == "xor" | block == "some") %>%
  ggplot(., aes(x = prior_class, y = response_centered)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_smooth(method="lm") +
  ylab("Inference strength ratings") +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Inference strength ratings by-predictor")

```

Plot the empirical predictor ratings against target inference ratings:
```{r}
d_critical_zScored_wide <- d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )

d_critical_zScored_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  ylab("Inference strength ratings") +
  geom_smooth(method = "lm") +
  facet_wrap(~main_type, ncol = 1) -> p.rel.z

d_critical_zScored_wide %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp.z

d_critical_zScored_wide %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri.z

gridExtra::grid.arrange(p.rel.z, p.comp.z, p.pri.z, ncol = 3) 
```
Plot by-item trends between predictor ratings and target inference strength ratings (relevance / competence ratings with utterance are omitted):
```{r, fig.height=35, fig.width=15}
# make a versatile wide representation of the critical data
d_critical_predictorsXtarget <- d_critical_zScored_wide %>% 
  pivot_longer(
    cols = c(relevance, prior, competence),
    names_to = "predictors",
    values_to = "predictor_response"
  ) 
d_critical_predictorsXtarget %>%
  ggplot(., aes(x = predictor_response, y = target)) +
  geom_point() +
  facet_wrap(title~predictors, ncol= 6) +
  xlab("Response to predictor questions") +
  ylab("Inference strength rating") +
  ggtitle("By-item Inference Strength Ratings vs. Predictor Ratings (left to right)")

```

Plot inference strength ratings against predictor ratings, by-item, using by-item mean ratings across participants.
```{r, fig.height=10, fig.width=8}
d_critical_predictorsXtarget_summary <- d_critical_predictorsXtarget %>% 
  group_by(title, predictors, main_type) %>%
  summarise(mean_target = mean(target),
            mean_predictor = mean(predictor_response), 
            sd_target = sd(target),
            sd_predictor = sd(predictor_response))

d_critical_predictorsXtarget_summary %>%
  ggplot(., aes(x = mean_predictor, y = mean_target)) +
  geom_point() +
  geom_text(aes(label=title), size = 2.5) +
  geom_smooth(method = "lm") +
  ylab("Mean inference strength rating") +
  xlab("Mean predictor ratings") + 
  facet_wrap(main_type~predictors)
```

Correlation plots of z-scored data:
``` {r}

# correlation plot for "some"
GGally::ggpairs(
  filter(d_critical_zScored_wide, main_type == "some") %>% ungroup() %>%  
    select(prior, competence, relevance, target)
)

# correlation plot for "or"
GGally::ggpairs(
  filter(d_critical_zScored_wide, main_type == "xor") %>%  ungroup() %>%
    select(prior, competence, relevance, target)
  )

```

## Stats

Our primary analysis is also conducted on z-scored data. 
First, maximal models on the ratings provided by the subjects are run:
```{r, results='hide'}
# xor, maximal model with interactions and maximal REs on z-scored data
model_xor_zScored <- brm(
  bf(target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored)
```

Plot correlations of main effects
```{r}
bayesplot::mcmc_pairs(model_xor_zScored, pars = c("b_prior", "b_relevance", "b_competence"))
```

```{r, results='hide'}
# some, maximal model with interactions and maximal REs
model_some_zScored <- brm(
  bf(target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored)
```

Plot correlations of main effects
```{r}
bayesplot::mcmc_pairs(model_some_zScored, pars = c("b_prior", "b_relevance", "b_competence"))
```

#### Smaller models with z-scored data to check for collinearity

Models with only one predictor: 
```{r, results='hide'}
# xor
model_xor_zScored_prior <- brm(
  bf(target ~ prior + 
    (1 + prior | submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored_prior)
```

```{r, results='hide'}
# xor
model_xor_zScored_comp <- brm(
  bf(target ~ competence + 
    (1 + competence | submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored_comp)
```

```{r, results='hide'}
# xor
model_xor_zScored_rel <- brm(
  bf(target ~ relevance + 
    (1 + relevance | submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored_rel)
```

```{r, results='hide'}
# some
model_some_zScored_rel <- brm(
  bf(target ~ relevance + 
    (1 + relevance | submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored_rel)
```

```{r, results='hide'}
# some
model_some_zScored_prior <- brm(
  bf(target ~ prior + 
    (1 + prior | submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored_prior)
```

```{r, results='hide'}
# some
model_some_zScored_comp <- brm(
  bf(target ~ competence + 
    (1 + competence | submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored_comp)
```

With two predictors:
```{r, results='hide'}
# xor
model_xor_zScored_pri_comp <- brm(
  bf(target ~ prior*competence + 
    (1 + prior*competence || submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored_pri_comp)
```

```{r}
bayesplot::mcmc_pairs(model_xor_zScored_pri_comp, pars = c("b_prior", "b_competence"))
```

```{r, results='hide'}
# xor
model_xor_zScored_pri_rel <- brm(
  bf(target ~ prior*relevance + 
    (1 + prior*relevance || submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored_pri_rel)
```

```{r}
bayesplot::mcmc_pairs(model_xor_zScored_pri_rel, pars = c("b_prior", "b_relevance"))
```

```{r, results='hide'}
# xor
model_xor_zScored_rel_comp <- brm(
  bf(target ~ relevance*competence + 
    (1 + relevance*competence || submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored_rel_comp)
```

```{r}
bayesplot::mcmc_pairs(model_xor_zScored_rel_comp, pars = c("b_relevance", "b_competence"))
```

```{r, results='hide'}
# some
model_some_zScored_pri_comp <- brm(
  bf(target ~ prior*competence + 
    (1 + prior*competence || submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored_pri_comp)
```

```{r}
bayesplot::mcmc_pairs(model_some_zScored_pri_comp, pars = c("b_prior", "b_competence"))
```

```{r, results='hide'}
# some
model_some_zScored_pri_rel <- brm(
  bf(target ~ prior*relevance + 
    (1 + prior*relevance || submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored_pri_rel)
```

```{r}
bayesplot::mcmc_pairs(model_some_zScored_pri_rel, pars = c("b_prior", "b_relevance"))
```

```{r, results='hide'}
# some
model_some_zScored_rel_comp <- brm(
  bf(target ~ relevance*competence + 
    (1 + relevance*competence || submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored_rel_comp)
```

```{r}
bayesplot::mcmc_pairs(model_some_zScored_rel_comp, pars = c("b_relevance", "b_competence"))
```

#### Stats with dichotomized predictors

Maximal model on z-scored inference ratings, using anticipated binary categorization of items as predictors 
```{r, results='hide'}
d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered, prior_class, class_condition) %>% 
  unique() %>% 
  pivot_wider(
    names_from = class_condition, 
    values_from = prior_class
  )  %>%
  mutate(relevance = factor(relevance, levels = c(1, 0)),
         competence = factor(competence, levels = c(1, 0)),
         prior = factor(prior, levels = c(1, 0))) %>% 
  filter(block_extended == "target") -> d_critical_wide_cat_zScored
# sum code predictors
# low prior: -1, high prior:: 1 
contrasts(d_critical_wide_cat_zScored$prior) <- contr.sum(2)
# low comp : -1, high comp : 1
contrasts(d_critical_wide_cat_zScored$competence) <- contr.sum(2)
# low rel: -1, high rel: 1
contrasts(d_critical_wide_cat_zScored$relevance) <- contr.sum(2)

# xor, maximal model with interactions and maximal REs
model_xor_cat_zScored <- brm(
  bf(response_centered ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_cat_zScored)
```

We test the following hypotheses for the categorical predictors models (identical for both or and some):

H1: For higher prior of the stronger alternative being true, the inference strength is lower

H2: For higher speaker comptence, the inference strength is higher

H3: For higher listener relevance, the inference strength is higher

```{r}
# xor
# test H1
hypothesis(model_xor_cat_zScored, "prior1 < 0")
# test H2
hypothesis(model_xor_cat_zScored, "competence1 > 0")
# test H3
hypothesis(model_xor_cat_zScored, "relevance1 > 0")
```

```{r, results='hide'}
model_some_cat_zScored <- brm(
  bf(response_centered ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_cat_zScored)
```

```{r}
# some
# test H1
hypothesis(model_some_cat_zScored, "prior1 < 0")
# test H2
hypothesis(model_some_cat_zScored, "competence1 > 0")
# test H3
hypothesis(model_some_cat_zScored, "relevance1 > 0")
```

#### Single predictor z-scored models with categorical predictors
For xor:
```{r, results="hide"}
model_xor_cat_zScored_pri <- brm(
  bf(response_centered ~ prior + 
    (1 + prior | submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_xor_cat_zScored_pri)
```

```{r, results="hide"}
model_xor_cat_zScored_comp <- brm(
  bf(response_centered ~ competence + 
    (1 + competence | submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_xor_cat_zScored_comp)
```

```{r, results="hide"}
model_xor_cat_zScored_rel <- brm(
  bf(response_centered ~ relevance + 
    (1 + relevance | submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_xor_cat_zScored_rel)
```

For some:
```{r, results="hide"}
model_some_cat_zScored_pri <- brm(
  bf(response_centered ~ prior + 
    (1 + prior | submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_some_cat_zScored_pri)
```

```{r, results="hide"}
model_some_cat_zScored_comp <- brm(
  bf(response_centered ~ competence + 
    (1 + competence | submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_some_cat_zScored_comp)
```

```{r, results="hide"}
model_some_cat_zScored_rel <- brm(
  bf(response_centered ~ relevance + 
    (1 + relevance | submission_id) +
    (1 | title),
    decomp = "QR"),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_some_cat_zScored_rel)
```

#### Tertiary analysis with by-item means (from paper)
Now we explore the original analysis with the by-item means:
```{r, results='hide'}
d_critical_zScored_wide_byItem_xor <- d_critical_zScored_wide %>% ungroup() %>%
  filter(main_type == "xor") %>%
  group_by(title) %>%
  summarize(mean_response = mean(target),
            mean_prior = mean(prior),
            mean_relevance = mean(relevance),
            mean_competence = mean(competence))

d_critical_zScored_wide_byItem_some <- d_critical_zScored_wide %>% ungroup() %>%
  filter(main_type == "some") %>%
  group_by(title) %>%
  summarize(mean_response = mean(target),
            mean_prior = mean(prior),
            mean_relevance = mean(relevance),
            mean_competence = mean(competence))

model_xor_means <- brm(bf(mean_response ~ mean_prior + mean_relevance + mean_competence,
                          decomp="QR"),
                       data = d_critical_zScored_wide_byItem_xor,
                       family = "gaussian",
                       cores = 4,
                       iter = 3000
                       )

model_some_means <- brm(bf(mean_response ~ mean_prior + mean_relevance + mean_competence,
                          decomp="QR"),
                       data = d_critical_zScored_wide_byItem_some,
                       family = "gaussian",
                       cores = 4,
                       iter = 3000
                       )
```

```{r}
summary(model_xor_means)
```
```{r}
summary(model_some_means)
```