### 2021-09-14 analysis/xor-some_final.Rmd
fixed slope coeff. computation, added hypotheses checks

---
title: "Xor & Some Final analysis"
author: "Polina Tsvilodub"
date: "8/31/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(brms)
library(tidyboot)
library(tidybayes)
```

## Read Data

First, the preprocessed and tidy wide pilot data is read and combined. It comprises data from N = 140 subjects after exclusions in total.
```{r, echo=FALSE}
pilot1 <- read_csv("./../data/pilots/pilot2_critical_zScore_wide_tidy.csv") %>%
  select(-competence_wUtt, -relevance_wUtt)
pilot2 <- read_csv("./../data/pilots/pilot3_critical_zScore_wide_tidy.csv")

d_both <- pilot1 %>% mutate(pilot = 1) %>% 
  rbind(., pilot2 %>% mutate(pilot = 2)
        )
# the data read here should be tidy, in correct format, bad workers should be excluded already
# z-scored data should be used and simulated!

# separate some and xor data!
d_both_some <- d_both %>% filter(main_type == "some") 
d_both_xor  <- d_both %>% filter(main_type == "xor")
```
Sanity check the combined data (as can be seen, z-scored data is used):
```{r, echo=FALSE, message=FALSE, warning=FALSE}
d_both %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("Inference strength ratings") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

d_both %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp

d_both %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri

gridExtra::grid.arrange(p.rel, p.comp, p.pri, ncol = 3) 
```

## Model

Next, fit the maximal model *across* experiments (i.e., *across SI trigger types*). For convergence reasons, correlation of random effects has to be set to 0. 
```{r, results='hide', message=FALSE, warning=FALSE}
# Full model 
d_both <- d_both %>% mutate(
  main_type = as.factor(main_type)
)
# some = 0, xor = 1
contrasts(d_both$main_type) 
# try to fit maximal model with brm
model_SI <- brm(target ~ prior*competence*relevance* main_type +
                   (1 + prior + competence + relevance + main_type || submission_id) +
                   (1 | title),
                 data = d_both,
                 control = list(adapt_delta = 0.95),
                 cores = 4,
                 iter = 3000)
```

```{r}
summary(model_SI)
```

## Extract conrtrasts of interest

Next, extract posterior contrasts of interest: effects of each predictor \{prior, competence and relevance\} for each trigger.

```{r}
model_SI %>% spread_draws(b_Intercept, b_prior, b_competence, b_relevance,
                          b_main_typexor, `b_prior:competence`,
                          `b_prior:relevance`, `b_competence:relevance`,
                          `b_prior:main_typexor`, `b_competence:main_typexor`,
                          `b_relevance:main_typexor`, `b_prior:competence:relevance`, 
                          `b_prior:competence:main_typexor`, `b_prior:relevance:main_typexor`,
                          `b_competence:relevance:main_typexor`, 
                          `b_prior:competence:relevance:main_typexor`) %>% 
  mutate(
    prior_xor = b_prior + `b_prior:main_typexor`,
    prior_some = b_prior,
    competence_xor = b_competence + `b_competence:main_typexor`,
    competence_some = b_competence,
    relevance_xor =  b_relevance + `b_relevance:main_typexor`,
    relevance_some = b_relevance
  ) -> model_SI_posteriors

# check P that the effects are positive / negative / no effect present
posterior_hypotheses <- model_SI_posteriors %>% 
  select(prior_xor, prior_some, 
         competence_xor, competence_some,
         relevance_xor, relevance_some) %>%
  gather(key, val) %>%
  group_by(key) %>% mutate(positive = mean(val > 0.05),
                           negative = mean(val < -0.05),
                           no = mean(val %>% between(-0.05, 0.05))) %>%
  summarise(positive_eff = mean(positive),
            negative_eff = mean(negative),
            no_eff = mean(no))
posterior_hypotheses
```

## Perform BF analysis

We want to address a conjunctive hypothesis, one triplet for each trigger word, namely:

1. the slope of 'competence' is positive
2. the slope of 'prior' is negative
3. the slope of 'relevance' is positive

We judge there to be strong evidence for positivity of a slope variable $\beta_X$, if the posterior probability $P(\beta_X > \delta \mid D)$ is at least $.95$, for $\delta = 0.05$ the parameter that defines our 'region of practical equivalence'.
A posterior odds ratio of at least $\frac{.95}{.05} = 19$ corresponds to a Bayes factor of at least 19 when prior odds are 1.

The binary test of conformity with the theoretical predictions therefore is:
```{r}
test_conjunction_of_all_hypotheses <-  function(posterior_hypotheses) {
  posterior_hypotheses %>% 
    mutate(hypothesis_true = case_when(
      key == 'competence_some' ~ positive_eff > 0.95,
      key == 'competence_xor'  ~ positive_eff > 0.95,
      key == 'prior_some' ~ negative_eff > 0.95,
      key == 'prior_xor'  ~ negative_eff > 0.95, 
      key == 'relevance_some' ~ positive_eff > 0.95,
      key == 'relevance_xor'  ~ positive_eff > 0.95
    )) %>% 
    pull(hypothesis_true) %>% all()
}
# applied to the pilot data
test_conjunction_of_all_hypotheses(posterior_hypotheses)
```

## Experimenting with multiple regression

```{r}
dummy_data <- tribble(
  ~main, ~REL, ~target,
  "some", 0, 4,
  "some", 1, 6,
  "xor", 0, 0,
  "xor", 1, 12
)
# check that the slope of "REL" for "xor" is 12
# this is calculated as betat_REL + beta_mainxor:REL
# importantly: the intercept adjustment beta_mainxor doesn't come in!
lm(target ~ main * REL, dummy_data)
```