### 2021-09-17 analysis/power/xor-some_bayesian-power.Rmd
refactor, add init simulation for Bayesian power analysis

---
title: "XOr-Some Bayesian Power analysis"
author: "Polina Tsvilodub"
date: "9/14/2021"
output: github_document
---

For the main xor-some experiment, we conduct a simulation-based Bayesian power analysis (used in a (loose) sense of determining the required sample size here). This power analysis aims at determining how many participants are required in order to detect a theoretically motivated conjunctive effect of the predictors prior, relevance and competence with a confidence of at least .8. The simulations are based on an assumed effect size $\beta = ±0.15$ for all predictors, and a region of practical equivalence (ROPE) $\delta = 0.05$ for judging evidence for the directionality of a coefficient given the data. The simulations are further based on the maximal desired model justified theoretically and by the design: $rating = trigger * prior * competence * relevance + REs$

In general, the power analysis proceeds as follows:

* the desired effect sizes of the critical predictors are set to $\mid 0.15\mid$ 
* hypothetical experimental data is simulated by drawing samples from a Normal distribution with the parameters described above 
* the desired maximal Bayesian regression model is computed on the simulated data 
* the N of subjects for which the data is simulated is increased iteratively in steps of 10 starting at 50
* for each N, the model is re-computed on the simulated samples for k=100 iterations (the more, the better the power estimate, but there are computational constraints)
* for each iteration $i_k$, the conjunctive hypothesis of interest is tested 
* the power for the given number of participants is calculated as the proportion of iterations for which the conjunctive hypothesis was credible (i.e., if the posterior probability $P(\beta_X > \delta \mid D)$ is at least $.95$, for $\delta = 0.05$ the parameter that defines our ROPE for the conjunction of all three predictor coefficients, for all six sub-hypotheses).
* we aim to determine the number of participants for which that proportion is at least .8.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidybayes)
library(brms)
```

## Simulate initial data
We assume an effect size of 0.15 for the predictors prior, competence and relevance, and, therefore, simulate data as coming from a Normal distribution with $\mu=±.15$ and $\sigma=.05$ (because this $\sigma$ would give us app. 95% of the data in the region we defined as a positive / negative effect). The target is simulated via $\mathcal{N}(\mu_t, 1)$, with $mu_t = \beta_0 + \beta_{pri} + \beta_{comp} + \beta_{rel} + \beta_{trigger} + interactions$. The coefficients for interactions and the trigger are assumed to be 0.

From each participant, we get four data points for xor and four data points for some, one data point per condition (prior X competence X relevance, mapped to triggers at random). For each condition (for each trigger), there are four stories for which the random intercept may vary.

Below, there is a toy example how data would be simulated for one subject (ignoring random effects for now).
```{r}
# ignore group-level effects for now; the effects will be sampled based on pilot data

# simulate toy data set for one subject 
d_init <-
  # create the 8 conditions
  tibble(prior = rep(c(0, 1), each = 4),
         comp = rep(c(0, 0, 1, 1), times = 2),
         rel = rep(c(0,1), times = 4),
         main_type = sample(rep(c("xor", "some"), times=4))) %>% 
  rowwise() %>%
  # sample effects based on condition
  mutate(
    subj_ID = 1, 
    intercept = 0,
    prior_obs = ifelse(prior == 0, rnorm(1, sd = 0.05), 
                       rnorm(1, -0.15, 0.05)),
    comp_obs = ifelse(comp == 0, rnorm(1, sd = 0.05), 
                       rnorm(1, 0.15, 0.05)),
    rel_obs = ifelse(rel == 0, rnorm(1, sd = 0.05), 
                       rnorm(1, 0.15, 0.05)),
    # sample response
    # what do we do about the interactions? can they safely be ignored?
    target = rnorm(1, 
                   mean = (intercept + prior_obs + comp_obs + rel_obs), 
                   sd = 1) # how to set / detrmine this properly?
  )
d_init
```

## Compute initial model
Next, the seed model is computed on the initial simulated dataset. This model will only be updated in further steps when the model is re-fit. 
```{r, eval=FALSE}
# fit model on simulated data for first time  
model_init <- brm(target ~ prior*comp*rel*main_type + 
                    (1 + prior + competence + relevance + main_type || submission_id) +
                   (1 | title),
                  data = d_init,
                  cores = 4,
                  contol = list(adapt_delta = 0.95),
                  iter = 3000)
model_init
```

## Build pipeline for simulating more data and fitting models

```{r, eval=FALSE}
# TODO
# just old stuff below
get_new_data <- function(d, N) {
    if(N == 0) {
      data <- d
    } else if (N > 47) {
      data <- add_predicted_draws(model=pilot_model, 
                                     newdata = pilot_data %>% 
                                       filter(workerid %in% sample(unique(pilot_data$workerid), N, replace = T)),
                                     n = 1) %>% 
                 mutate(workerid = paste(workerid, letters[1], sep = "_")) %>%
                   rbind(., d)
    } else {
      data <- add_predicted_draws(model=pilot_model, 
                                     newdata = pilot_data %>% 
                                       filter(workerid %in% sample(unique(pilot_data$workerid), N, replace = F)),
                                     n = 1) %>% 
                 mutate(workerid = paste(workerid, letters[1], sep = "_")) %>%
                   rbind(., d)
    }

  }

```

## Extract contrasts of interest

```{r}
# TODO
```

## Test H of interest

```{r}
# TODO
```