### 2021-10-17 analysis/power/xor-some_bayesian-power_500.Rmd
final 500 iter power simulation

---
title: "XOr-Some Bayesian Power analysis"
author: "Polina Tsvilodub"
date: "9/14/2021"
output: 
  github_document:
  pandoc_args: --webtex
---

For the main xor-some experiment, we conduct a simulation-based Bayesian power analysis (used in a (loose) sense of determining the required sample size here). This power analysis aims at determining how many participants are required in order to detect a theoretically motivated conjunctive effect of the predictors prior, relevance and competence with a confidence of at least .8. The simulations are based on an assumed effect size $\beta = 0.15$ for all predictors, and a region of practical equivalence (ROPE) $\delta = 0.05$ for judging evidence for the directionality of a coefficient given the data. The simulations are further based on the maximal desired model justified theoretically and by the design: `rating = trigger * prior * competence * relevance`

In general, the power analysis proceeds as follows:

* the desired effect sizes of the critical predictors are set to $0.15$ 
* hypothetical experimental data is simulated by drawing samples from a Normal distribution with the parameters described above 
* the desired maximal Bayesian regression model is computed on the simulated data 
* the N of subjects for which the data is simulated is increased iteratively in steps of 20 starting at 100 up to 340.
* for each N, the model is re-computed on the simulated samples for k=100 iterations (the more, the better the power estimate, but there are computational constraints)
* for each iteration $i_k$, the conjunctive hypothesis of interest is tested 
* the power for the given number of participants is calculated as the proportion of iterations for which the conjunctive hypothesis was credible (i.e., if the posterior probability $P(\beta_X > \delta \mid D)$ is at least $.95$, for $\delta = 0.05$ the parameter that defines our ROPE for the conjunction of all three predictor coefficients, for all six sub-hypotheses).
* we aim to determine the number of participants for which that proportion is at least .8.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidybayes)
library(brms)
```

## Simulate initial data
We assume an effect size of 0.15 for the predictors prior, competence and relevance. We simulate data as coming from a Normal distribution with $\mu=0$ and $\sigma=1$ and z -score it, to follow the experiemental procedure.

The target response variable is simulated via $\mathcal{N}(\mu_t, 1)$, with $\mu_t = 0.15 * ( \beta_\text{pri} + \beta_\text{comp} + \beta_\text{rel} + \beta_\text{trigger} ) + \text{interactions}$. The multiplication by 0.15 represent hereby the assumed effect size. The coefficients for interactions and the trigger effect are assumed to be 0.

From each participant, we get four data points for xor and four data points for some, one data point per condition (prior X competence X relevance, mapped to triggers at random). Therefore, we simulate 8 trials per subject.

```{r}
# create fake data
create_fake_data <- function(N = 100, seed = 1) {
  set.seed(seed)
  beta_rel  <- 0.15
  beta_comp <- 0.15
  beta_pri  <- 0.15
  map_df(
    1:N, 
    function(i) {
      tibble(
        subj    = rep(str_c("subj_",i), times = 8),
        trigger = rep(c('or', 'some'), times= 4), 
        rel     = rnorm(8),
        comp    = rnorm(8),
        pri     = rnorm(8)
      ) %>% 
        group_by(subj) %>% 
        mutate(
          rel     = (rel-mean(rel)) / sd(rel),
          comp    = (comp-mean(comp)) / sd(comp),
          pri     = (pri-mean(pri)) / sd(pri)
        ) %>% 
        ungroup() %>% 
        mutate(
          target  = rnorm(8, beta_rel * rel + beta_comp * comp + beta_pri * pri)
        )
    }
  )
}

```

## Compute initial model
Next, the seed model is computed on the initial simulated dataset. This model will only be updated in further steps when the model is re-fit. 
```{r}
# create initial dataset 
initial_fake_data <- create_fake_data(N=100, seed = 1234)

# initial fit
fit_initial <- brm(bf(target ~ rel * comp * pri,
                      decomp = "QR"), 
                   initial_fake_data)

```

## Build pipeline for simulating more data and fitting models

Build helper function for simulating and computing regression model for k iterations (tracked as seeds). Also extract contrasts of interest here.
```{r, message = FALSE, warning=FALSE, results='hide'}
# main results
start_N <- 100
step <- 20
stop_N <-340
seed <- 500

sim_data_fit <- function(N, k){
 results <- map_dbl(
  1:k, 
  function(i) {
    fake_data <- create_fake_data(N=N, seed = i)
    fit <- update(fit_initial, newdata = fake_data, seed = i, cores = 4)
    lower_bounds <- rbind(
      aida::summarize_sample_vector(brms::posterior_samples(fit)[,'b_rel'], "rel"),
      aida::summarize_sample_vector(brms::posterior_samples(fit)[,'b_comp'], "comp"),
      aida::summarize_sample_vector(brms::posterior_samples(fit)[,'b_pri'], "pri")
    ) %>% pull("|95%")
    min(lower_bounds)
  }
 )
 # create path for file
 stream_out <- paste("./results/power_sim_N", N, "_iter", k, ".csv", sep="")
 # create tibble for intermediate saving
 tibble(
   seed = 1:k, lower_bounds = results
 ) %>%
   write_csv(., stream_out)
}

sim_power <- function(seed, start_N, step, stop_N) {
  # iterate over different Ns of subjects
  sim_full <- tibble( N = seq(start_N, stop_N, step)) %>%
    mutate(
      tidy = map(N, sim_data_fit, k = seed)
    ) %>%
    unnest(tidy)
  write_csv(sim_full, paste("./results/power_sim_full_N", start_N, "-N", stop_N, "_iter", seed, ".csv", sep=""), append = T, col_names = T)
}

power_df_500 <- sim_power(seed, start_N, step, stop_N)

```

Below, the results are summarized by simulated N of subjects to obtain the power estimates (i.e., proportion of simulation runs where all three effects were >0.05).
```{r}
# analyse power summary
power_results_500 <- power_df_500 %>% group_by(N) %>% 
  summarise(power = mean(lower_bounds > 0.05))

write_csv(power_results_500, "./results/power_sim_500_results_summary.csv")

power_results_500
```

Finally, power results are visualized as a function of the N of subjects:
```{r}
power_results_500 %>%
  ggplot(., aes(x = N, y = power)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0.8, alpha = 0.7, linetype = 2) +
  ggtitle("Power by N of subjects") 
ggsave("./results/power_sim_results_plot.png")
```