### 2021-10-18 analysis/prereg/02_xor-some_prereg_plots.Rmd

---
title: 'Xor-Some Preregistered Study: Visualizations'
author: "Polina Tsvilodub"
date: "10/18/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidyboot)
```

Read data created in preprocessing script:
```{r}
d_critical_zScored_wide <- read_csv("./../../data/results_prereg_tidy_zScored_wide.csv")
d_critical_zScored <- read_csv("./../../data/results_prereg_tidy_zScored_long.csv")
d_test_raw <- read_csv("./../../results_prereg_raw_test.csv")
```

## Plots

Comprehension question performance:
```{r}
d_test_ci <- d_test %>% group_by(test_condition) %>% 
  tidyboot_mean(column = response)

d_test %>% 
  ggplot(., aes(x = test_condition, y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
  geom_point(data = d_test_ci, aes(x = test_condition, y = mean), color = "red", 
             size = 4) +
  facet_wrap(~main_type)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
d_critical_zScored_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  ylab("Inference strength ratings") +
  geom_smooth(method = "lm") +
  facet_wrap(~main_type, ncol = 1) -> p.rel.z

d_critical_zScored_wide %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp.z

d_critical_zScored_wide %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri.z

gridExtra::grid.arrange(p.rel.z, p.comp.z, p.pri.z, ncol = 3) 
```

```{r}
d_critical_zScored %>% 
  filter(block == "xor" | block == "some") %>%
  ggplot(., aes(x = prior_class, y = response_centered)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_smooth(method="lm") +
  ylab("Inference strength ratings") +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Inference strength ratings by-predictor")

```