### 2021-11-07 analysis/main/03_xor-some_prereg_analysis_final.Rmd

---
title: "Xor-Some Preregistered Study: Analyses"
author: "Polina Tsvilodub"
date: "8/31/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(brms)
library(tidyboot)
library(tidybayes)
```

Read data created in preprocessing script:
```{r, message=FALSE, warning=FALSE, results='hide'}
d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
```

## Main analysis

Next, fit the maximal model *across* experiments (i.e., *across SI trigger types*). For convergence reasons, correlation of random effects has to be set to 0. 
```{r, results='hide', message=FALSE, warning=FALSE}
# Full model 
d_critical_zScored_wide <- d_critical_zScored_wide %>% mutate(
  main_type = as.factor(main_type)
)
# some = 0, xor = 1
contrasts(d_critical_zScored_wide$main_type)

# set priors
priors <- set_prior("student_t(1, 0, 2)", class = "b")

model_SI <- brm(target ~ prior*competence*relevance* main_type +
                   (1 + prior + competence + relevance + main_type || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.95),
                 cores = 4,
                 iter = 3000)
```

```{r}
summary(model_SI)
```

Next, extract posterior contrasts of interest: effects of each predictor \{prior, competence and relevance\} for each trigger.

```{r}
model_SI %>% spread_draws(b_Intercept, b_prior, b_competence, b_relevance,
                          b_main_typexor, `b_prior:competence`,
                          `b_prior:relevance`, `b_competence:relevance`,
                          `b_prior:main_typexor`, `b_competence:main_typexor`,
                          `b_relevance:main_typexor`, `b_prior:competence:relevance`, 
                          `b_prior:competence:main_typexor`, `b_prior:relevance:main_typexor`,
                          `b_competence:relevance:main_typexor`, 
                          `b_prior:competence:relevance:main_typexor`) %>% 
  mutate(
    prior_xor = b_prior + `b_prior:main_typexor`,
    prior_some = b_prior,
    competence_xor = b_competence + `b_competence:main_typexor`,
    competence_some = b_competence,
    relevance_xor =  b_relevance + `b_relevance:main_typexor`,
    relevance_some = b_relevance
  ) -> model_SI_posteriors

# check P that the effects are positive / negative / no effect present
posterior_hypotheses <- model_SI_posteriors %>% 
  select(prior_xor, prior_some, 
         competence_xor, competence_some,
         relevance_xor, relevance_some) %>%
  gather(key, val) %>%
  group_by(key) %>% mutate(positive = mean(val > 0.05),
                           negative = mean(val < -0.05),
                           no = mean(val %>% between(-0.05, 0.05))) %>%
  summarise(positive_eff = mean(positive),
            negative_eff = mean(negative),
            no_eff = mean(no))
posterior_hypotheses
```

#### Perform BF analysis

We want to address a conjunctive hypothesis, one triplet for each trigger word, namely:

1. the slope of 'competence' is positive
2. the slope of 'prior' is negative
3. the slope of 'relevance' is positive

We judge there to be strong evidence for positivity of a slope variable $\beta_X$, if the posterior probability $P(\beta_X > \delta \mid D)$ is at least $.95$, for $\delta = 0.05$ the parameter that defines our 'region of practical equivalence'.
A posterior odds ratio of at least $\frac{.95}{.05} = 19$ corresponds to a Bayes factor of at least 19 when prior odds are 1.

The binary test of conformity with the theoretical predictions therefore is:
```{r}
test_conjunction_of_all_hypotheses <-  function(posterior_hypotheses) {
  posterior_hypotheses %>% 
    mutate(hypothesis_true = case_when(
      key == 'competence_some' ~ positive_eff > 0.95,
      key == 'competence_xor'  ~ positive_eff > 0.95,
      key == 'prior_some' ~ negative_eff > 0.95,
      key == 'prior_xor'  ~ negative_eff > 0.95, 
      key == 'relevance_some' ~ positive_eff > 0.95,
      key == 'relevance_xor'  ~ positive_eff > 0.95
    )) %>% 
    pull(hypothesis_true) %>% all()
}
# applied to the pilot data
test_conjunction_of_all_hypotheses(posterior_hypotheses)
```
### Plot posterior distributions

```{r, fig.height = 6, fig.width=9}
model_SI_posteriors %>% select(prior_xor, prior_some,
                               competence_xor, competence_some,
                               relevance_xor, relevance_some) %>%
  pivot_longer(cols = everything(), names_to = "effect", values_to = "value") -> model_SI_posteriors_long

model_SI_posteriors_long %>%
  ggplot(.) + # 
  stat_halfeye(aes(y = effect, x = value, 
                   fill = stat(case_when(x %>% between(-0.05, 0.05) ~ "no effect", x < -0.05 ~ "negative effect", x > 0.05 ~ "positive effect"))
                   ), 
               alpha = .7) +
  geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
  scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) +
  ggtitle("Posterior effect distributions. Points indicate posterior means\nThick lines indicate 66% CrIs, thin lines 95% CrIs.")

```

## Exploratory analyses

Model with binary predictors:
```{r, results='hide', message=FALSE, warning=FALSE}
d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered, prior_class, class_condition) %>% 
  unique() %>% 
  pivot_wider(
    names_from = class_condition, 
    values_from = prior_class
  )  %>%
  mutate(relevance = factor(relevance, levels = c(1, 0)),
         competence = factor(competence, levels = c(1, 0)),
         prior = factor(prior, levels = c(1, 0)),
         main_type = as.factor(main_type)) %>% 
  filter(block_extended == "target") -> d_critical_wide_cat_zScored
# sum code predictors
# low prior: -1, high prior:: 1 
contrasts(d_critical_wide_cat_zScored$prior) <- contr.sum(2)
# low comp : -1, high comp : 1
contrasts(d_critical_wide_cat_zScored$competence) <- contr.sum(2)
# low rel: -1, high rel: 1
contrasts(d_critical_wide_cat_zScored$relevance) <- contr.sum(2)
# leave main type dummy coded
contrasts(d_critical_wide_cat_zScored$main_type)

# xor, maximal model with interactions and maximal REs
model_xor_cat_zScored <- brm(
  response_centered ~ prior*competence*relevance*main_type + 
    (1 + prior + competence + relevance + main_type || submission_id) +
    (1 | title),
  data = d_critical_wide_cat_zScored,
  prior = priors,
  sample_prior = T,
  family = "gaussian",
  cores = 4,
  iter = 3000
)

```
```{r}
summary(model_xor_cat_zScored)
```
Now extract posterior contrasts of interest from categorical model: effects of each predictor \{prior, competence and relevance\} for each trigger.

```{r}
model_xor_cat_zScored %>% spread_draws(b_Intercept, b_prior1, b_competence1, b_relevance1,
                          b_main_typexor, `b_prior1:competence1`,
                          `b_prior1:relevance1`, `b_competence1:relevance1`,
                          `b_prior1:main_typexor`, `b_competence1:main_typexor`,
                          `b_relevance1:main_typexor`, `b_prior1:competence1:relevance1`, 
                          `b_prior1:competence1:main_typexor`, `b_prior1:relevance1:main_typexor`,
                          `b_competence1:relevance1:main_typexor`, 
                          `b_prior1:competence1:relevance1:main_typexor`) %>% 
  mutate(
    prior_xor = 2 * b_prior1 + 2* `b_prior1:main_typexor`,
    prior_some = 2 * b_prior1,
    competence_xor = 2 * b_competence1 + 2* `b_competence1:main_typexor`,
    competence_some = 2 * b_competence1,
    relevance_xor =  2 * b_relevance1 + 2* `b_relevance1:main_typexor`,
    relevance_some = 2* b_relevance1
  ) -> model_xor_cat_zScored_posteriors

# check P that the effects are positive / negative / no effect present
posterior_hypotheses_cat <- model_xor_cat_zScored_posteriors %>% 
  select(prior_xor, prior_some, 
         competence_xor, competence_some,
         relevance_xor, relevance_some) %>%
  gather(key, val) %>%
  group_by(key) %>% mutate(positive = mean(val > 0.05),
                           negative = mean(val < -0.05),
                           no = mean(val %>% between(-0.05, 0.05))) %>%
  summarise(positive_eff = mean(positive),
            negative_eff = mean(negative),
            no_eff = mean(no))
posterior_hypotheses_cat
```
```{r}
test_conjunction_of_all_hypotheses(posterior_hypotheses_cat)
```

```{r, fig.height = 6, fig.width=9}
model_xor_cat_zScored_posteriors %>% select(prior_xor, prior_some,
                               competence_xor, competence_some,
                               relevance_xor, relevance_some) %>%
  pivot_longer(cols = everything(), names_to = "effect", values_to = "value") -> model_xor_cat_posteriors_long

model_xor_cat_posteriors_long %>%
  ggplot(.) + # 
  stat_halfeye(aes(y = effect, x = value, 
                   fill = stat(case_when(x %>% between(-0.05, 0.05) ~ "no effect", x < -0.05 ~ "negative effect", x > 0.05 ~ "positive effect"))
                   ), 
               alpha = .7) +
  geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
  scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) +
  ggtitle("Posterior effect distributions. Points indicate posterior means\nThick lines indicate 66% CrIs, thin lines 95% CrIs.")

```

Model with aggregated by-story ratings:
```{r, results='hide', message=FALSE, warning=FALSE}
d_critical_zScored_wide_byItem <- d_critical_zScored_wide %>% ungroup() %>%
  group_by(title, main_type) %>%
  summarize(mean_response = mean(target),
            mean_prior = mean(prior),
            mean_relevance = mean(relevance),
            mean_competence = mean(competence))

model_means <- brm(mean_response ~ mean_prior * mean_relevance * mean_competence * main_type,
                       data = d_critical_zScored_wide_byItem,
                       prior = priors,
                       sample_prior = T,
                       family = "gaussian",
                       cores = 4,
                       iter = 3000
                       )
```

```{r}
summary(model_means)
```

Extract posterior contrasts of interest from by-story mean based model: effects of each predictor \{prior, competence and relevance\} for each trigger.

```{r}
model_means %>% spread_draws(b_Intercept, b_mean_prior, b_mean_competence, b_mean_relevance,
                          b_main_typexor, `b_mean_prior:mean_competence`,
                          `b_mean_prior:mean_relevance`, `b_mean_relevance:mean_competence`,
                          `b_mean_prior:main_typexor`, `b_mean_competence:main_typexor`,
                          `b_mean_relevance:main_typexor`,
                          `b_mean_prior:mean_relevance:mean_competence`, 
                          `b_mean_prior:mean_competence:main_typexor`,
                          `b_mean_prior:mean_relevance:main_typexor`,
                          `b_mean_relevance:mean_competence:main_typexor`, 
                          `b_mean_prior:mean_relevance:mean_competence:main_typexor`) %>% 
  mutate(
    prior_xor = b_mean_prior + `b_mean_prior:main_typexor`,
    prior_some = b_mean_prior,
    competence_xor = b_mean_competence + `b_mean_competence:main_typexor`,
    competence_some = b_mean_competence,
    relevance_xor =  b_mean_relevance + `b_mean_relevance:main_typexor`,
    relevance_some = b_mean_relevance
  ) -> model_means_posteriors

# check P that the effects are positive / negative / no effect present
posterior_hypotheses_means <- model_means_posteriors %>% 
  select(prior_xor, prior_some, 
         competence_xor, competence_some,
         relevance_xor, relevance_some) %>%
  gather(key, val) %>%
  group_by(key) %>% mutate(positive = mean(val > 0.05),
                           negative = mean(val < -0.05),
                           no = mean(val %>% between(-0.05, 0.05))) %>%
  summarise(positive_eff = mean(positive),
            negative_eff = mean(negative),
            no_eff = mean(no))
posterior_hypotheses_means
```

```{r}
test_conjunction_of_all_hypotheses(posterior_hypotheses_means)
```

```{r, fig.height = 6, fig.width=9}
model_means_posteriors %>% select(prior_xor, prior_some,
                               competence_xor, competence_some,
                               relevance_xor, relevance_some) %>%
  pivot_longer(cols = everything(), names_to = "effect", values_to = "value") -> model_mean_posteriors_long

model_mean_posteriors_long %>%
  ggplot(.) + # 
  stat_halfeye(aes(y = effect, x = value, 
                   fill = stat(case_when(x %>% between(-0.05, 0.05) ~ "no effect", x < -0.05 ~ "negative effect", x > 0.05 ~ "positive effect"))
                   ), 
               alpha = .7) +
  geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
  scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) +
  ggtitle("Posterior effect distributions. Points indicate posterior means\nThick lines indicate 66% CrIs, thin lines 95% CrIs.")

```