### 2021-11-24 writing/paper/paper_draft.tex
fix typos

\documentclass{sp}

% The \pdf* commands provide metadata for the PDF output.
% Do not use LaTeX style / commands like \emph{} inside these.
\pdfauthor{Author Full Name(s)}
\pdftitle{The role of priors, competence and relevance for the interpretation of disjunction}
\pdfkeywords{exclusive disjunction, scalar implicature, pragmatics}

% Optional short title inside square brackets, for the running headers.
% If no short title is given, no title appears in the headers.
\title[Exclusive disjunction]{The role of priors, competence and relevance for the interpretation of disjunction%
  \thanks{We thank \ldots}}

% Optional short author inside square brackets, for the running headers.
% If no short author is given, no authors print in the headers.
\author[]{% As many authors as you like, each separated by \AND.
  \spauthor{Michael Franke \\ \institute{Institute1}} \AND
  \spauthor{Bob van Tiel \\ \institute{Institute2}} \AND
  \spauthor{Polina Tsvilodub \\ \institute{Institute3}}}

\begin{document}

\maketitle

\begin{abstract}

  Sentences containing disjunctions like ``Donald ate a pretzel or a donut'' are often taken as conveying that Donald ate a pretzel or a donut, but not both. That is, the listener might infer an \textit{exclusive reading of the disjunction}. Current standard accounts of disjunction interpretation posit that the exclusive interpretation of disjunctions is an instance of \textit{scalar implicature}. Prior work suggests that, among others, three factors influence the robustness of scalar implicatures: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability} that the stronger alternative is true. While their influence was investigated experimentally for the interpretation of ``some'', the evidence is less conclusive in the case of disjunction interpretation. Therefore, we compare experimentally the interpretation of ``some'' and ``or'', while manipulating the three factors with respect to the stronger alternatives “all”, and “and”, respectively. Our results indicate that competence has a consistent effect of the robustness of the implicature generation for both triggers, while prior and relevance exhibit different, more intricate, effects varying between the triggers.

\end{abstract}

\begin{keywords}
 exclusive disjunction, scalar implicature, pragmatics
\end{keywords}

\section{Introduction}

General intuition, introduction \citep[e.g.,][]{horn1972semantic}.

\subsection{Scalar implicature}

SI accounts and treatments of ``some'' and ``or''.

\textbf{Related experimental/modelling work: } 
Some: \citep{degen2015investigating}, 
Or: \citep{Li2021}

\section{Relevance, competence and prior pribability}
Literature on the effects of these factors on SIs \citep{sperber1986relevance, goodman2013knowledge, degen2015wonky}. 
\subsection{Relevance}

\subsection{Competence}

\subsection{Prior probability}

\subsection{Predictions}
Prior research has shown that the robustness of implicatures from ``some'' to ``some, but not all''  predicted by the standrard scalar implicature (SI) account is influenced by three factors. First, high contextual relevance of the stronger alternative ``all'' to the listener is predicted to render more robust scalar implicatures. Second, high speaker competence about the truth of the stronger alternative is predicted to lead to more robust scalar implicature generation. Lastly, high prior probability of the stronger ``all'' being true is assumed to \textit{decrease} the scalar implicature rate.
If the currently standard scalar implicature account also holds for the interpretation of disjunctions, analogous predictions for the robustness of the implicature from ``A or B'' to ``A or B, but not both'' can be derived. In case of disjunction, the relevant stronger alternative is ``and''.

We hypothesize that all three factors influence the robustness of scalar implicatures for both triggers. That is, we expect a higher rate of endorsements for sentences containing the upper-bounded (exclusive, in case of ``or'') paraphrases of the respective triggers when the three factors are intuitively judged to be high or low, as described above. We, therefore, test six hypotheses, one concerning the effect of each factor for each trigger. 
Simultaneously,  we also address the \textit{identity hypothesis} that exclusive readings of disjunctions are a scalar implicature just like the ``some → some but not all'' case. If the identity hypothesis is correct, we expect that whichever factors positively or negatively affect the implicature readings of ``some'', exactly these factors also affect the exclusive readings of ``or''. In other words, the prediction of the identity hypothesis could still be confirmed even if the hypotheses related to the effects of relevance, competence and prior are not borne out, as long as the observed results are the same for ``some'' and ``or''. To investigate this, we directly compare the interpretation of the two triggers within one experiment, described in the following section.

\section{Experiment}

In this experiment we directly compare the interpretation of ``some'' and ``or'', so as to directly explore if the predictions generated by the scalar implicature account are borne out comparably for both triggers. Specifically, we manipulate the \textit{contextual relevance of the stronger alternative to the listener}, the \textit{speaker's competence about the truth of ``all'' / ``and''}, and the \textit{prior probability of ``all'' / ``and'' being true}, to investigate the robustness of an exclusive reading of ``or'' (``A or B, but not both'') under these manipulations, in direct comparison to the upper-bounded reading of “some” (``some but not all'').

To operationalize these factors, we designed context stories (vignettes) which were judged to score either high or low with respect to each factor of interest (high/low prior probability of the more informative alternative $\times$ high/low speaker competence $\times$ high/low contextual relevance of the alternative). Combining these dichotomozed manipulations resulted in eight unique conditions.

On critical trials, participants were asked to rate four sentences, one per factor and one containing an upper-bounded (exclusive) paraphrase of the trigger. The sentences elicited judgements how likely it is that each factor was high, and how likely the critical scalar implicature was true. The context stories and sentences manipulating the factors of interest were designed in an analogous fashion for both triggers and varied within-subjects, so the following descriptions apply for both triggers.

If our factor hypotheses are true, on average, we expect higher likelihood ratings for the statements eliciting relevance, competence and prior judgements given items in the respective high conditions. Critically, we expect higher likelihood ratings for the scalar implicature, if the relevance and comptenece in the item were judged as high and the prior of the alternative as low.
%Participants were then asked to rate statements eliciting four judgements using a slider bar: the relevance of the “all”/“and”-alternative, the competence of the speaker about the alternative, the prior probability of ”all” / “and” being true, and, crucially, the likelihood of the exclusive reading of “or” or of the upper-bounded reading of “some”.

\subsection{Procedure and Materials}
This study is a two-by-two-by-two-by-two within-subjects rating task (relevance $\times$ competence $\times$ prior $\times$ trigger), conducted as a web-based experiment. The study proceeds as follows.

On each trial, participants read a background story, followed by a sentence presented in a blue box which is meant to elicit the likelihood rating. These sentences were of the form ``It is important to X to know whether Y'' (relevance rating), ``Z knows whether Y'', ``If A, then B'' / ``If B , then A'' (disjunction prior rating), ``If some Y, then all Y'' (some prior rating), ``From what Z said we may conclude that W'', where X is the listener, Z is the speaker and Y is the target event in the background story, and W is the event under the upper bounded / exclusive reading. A and B are the disjuncts in case of disjunction. For the inference strength elicitation, participants additionally read a critical utterance of the form ``Z says to X: Y'', containing the trigger (``some'' / ``or''), presented below the background story in a red box.

Below, participants read the task saying ``How likely is it that the statement is true given the story?'', followed by a slider rating bar labelled ``certainly false'' (left) and ``certainly true'' (right). The continuous ratings were converted to an integer rating between 0 and 100 in the background. For each critical vignette, participants completed a block of 10 or 11 trials, consisting of a trial with a comprehension question, trials with statements eliciting relevance, comptenece and prior ratings each (the last factor was elicited with two statements for the trigger ``or''), followed by three more comprehension questions, and the critical implicature strength elicitation trial. 
Comprehension trials were visually identical to the critical trials, but the statement to be rated only referred to the content of the background stories, and were designed so as to be either clearly true, clearly false or uncertain given the background story. 
The four comprehension statements for one vignette were sampled at random from six possible statements (two per type true / false / uncertain).

 We designed 32 different stories per trigger type (``some''~vs.~``or''), resulting in a total of 64 stories. Each subject sees eight randomly sampled stories (one per prior $\times$ competence $\times$ relevance condition out of four possible stories) such that they see four ``or'' and four ``some'' stories in randomized order. The assignment of conditions to the triggers is randomized between-subjects.

At the beginning, participants were welcomed to the experiment, read instructions and first saw three labelled example trials showing the working of the slider. They read a simple story about a person shopping. They then saw a question about the story that was clearly false (expected rating: 0), clearly true (expected rating: 100) and implied to be rather true (expected rating: 50-100), given the background story. These expected responses were labelled and explained.  
The main part of the experiment consisted of eight critical stories, randomly shuffled with eight attention check stories. The attention checks consisted of one trial, visually matching critical trials. On these trials, the vignettes contained a statement which clearly wrote out what participants were supposed to answer (e.g., ``Please move the slider maximally left'') in an area of text that participants had to read to complete the trial. Participants who failed more than two of these attention checks were excluded from the analysis. 
%Each critical story block has the following structure: Participants read a background story, followed by a comprehension statement which is either clearly true, clearly false, or uncertain (sampled at random for each story) which they are asked to rate using the slider. Then, they read statements presented in a box colored blue associated with the factors of interest and rate it using a slider bar. The factors of interest are relevance, competence and prior, in randomized order within-participants. In the “or”-condition, the prior statement consists of two symmetrical conditional sentences (see example below). Then, participants answer another three comprehension questions. These are followed by the critical utterance containing “or” / “some” appearing in a box colored red. Below, the target statement gathering the likelihood of the exclusive disjunction / the upper-bounded reading of “some” (i.e., the likelihood of the implicature) appears. Again, they rate the target statement using the slider. The background story remains on the screen throughout the trials. 
After the study, participants could fill out a socio-demographic questionnaire.
The experiment can be viewed under \texttt{https://magpie-xor-some.netlify.app/}. The preregistration for the experiment can be found under \texttt{https://osf.io/v7zjp}.

\subsection{Participants}
We recruited 277 participants through the crowd-sourcing platform Prolific. Participants were restricted to those whose first language included English, who previously took part in at least five other Prolific studies, and whose approval rate was at least 0.9, according to prescreening criteria of the platform. Participants took 20 minutes on average to complete the study and were compensated \pounds2.48 for their participation. Following our preregistered exclusion criteria, we excluded 15 participants for not indicating their native language, 3 participants for completing the study in under 8 minutes, 26 participants for failing more than 20\% of the attention checks, and 27 for failing more than 20\% of the comprehension questions. Due to a coding error of the comprehension questions, 206 participants were left post applying exclusion criteria, although the preregistered target sample size post exclusions was 200 subjects. We conducted a power analysis in order to determine the minimal sample size required for a power of 0.8 of finding evidence for the conjunction of all factor hypotheses (relevance, competence and prior) predicted by the SI account, finding that at least 180 subjects were required. 
In the following analyses, data from the 206 participants is analysed.\footnote{All analyses were also conducted on data from 200 subject, excluding the last six submissions. No noteworthy quantitative or qualitative differences of the results were observed.}

\section{Results}
Prior to conducting statistical analyses, we preprocess the data by standardizing (or z-scoring) the responses within each factor (relevance, comptence, prior) for each participant. This is done to counterbalance possible by-participant responsiveness to particular factors, while keeping an anlysis advantageous to the scalar implicature account, since the standardization is applied across triggers.

\subsection{Main analysis}
We analysed the results using a Bayesian linear mixed effects model, regressing the target implicature likelihood rating against the fixed effects of predictor ratings (i.e., the relevance, comptence and prior ratings elicited by the same participant for that vignette), the effect of trigger and their two-, three-, and four-way interactions. Maximal computationally tractable random effects structure justified by the experimental design was included: we included random intercepts and random slope effects for the main effects of trigger, relevance, competence and prior by-subject, as well as random intercepts by-vignette.\footnote{Model in R syntax style: \texttt{inference-rating $\sim$ trigger * relevance * competence * prior + (1 + trigger + relevance + competence + prior || subjectID) + (1 | vignette)}. For computational tractability reasons, the correlation of random effects by-subject was set to 0.} 
The categorical effect of trigger was dummy coded, using ``some'' as the reference level. For all regression coefficients we used a wide and uninformative prior given by a $t$-distribution with mean 0, standard deviation of 2 and 1 degree of freedom. The model was fitted using the R \texttt{brms} package \citep{burkner2017brms}.\footnote{The data and the analyses can be found under \texttt{https://github.com/magpie-ea/magpie-xor-experiment}}

We focus on six slope coefficients of the regression model, for the effects of relevance, competence and prior, respectively, once for ``some'' and once for ``or''. We check whether the posterior estimate of each effect is in either one of three intervals: (1) negative effect (<= -0.05), (2) no effect (between -0.05 and 0.05), or (3) positive effect (>= 0.05). 
We interpret the data as providing evidence in favor of an effect (positive, negative, no effect) if the posterior probability of the effect being true is >= 0.95 (i.e., 95\% of posterior samples are in the corresponding interval). In particular, we speak of evidence in favor of the scalar implicature account of the exclusive disjunction interpretation if the posterior probability of the prior effect being negative, the relevance and competence effects being positive is >= 0.95.

\begin{figure*}[t]
	\begin{center}
		\includegraphics[width=1\linewidth]{images/posterior-effects-main.png}
	\end{center}
	\vspace{-0.3cm}
	\caption{Main analysis results: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''. The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).}
	\label{posteriors-main}
\end{figure*}

Consistent with predictions, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true, as inidcated by the probability of the negative effect of prior being $P = 0.999$ (Fig.\ref{posteriors-main}, \texttt{prior\_some} line, orange color). Similarly, we find a clear positive effect for the effect of speaker competence ($P =  1$, Fig.\ref{posteriors-main}, \texttt{competence\_some} line, green color). However, we do not find any clear effects of relevance (Fig.\ref{posteriors-main}, \texttt{relevance\_some} line, split colors). If anything, the data supports the hypothesis that the effect is not negative.  
We hypothesize that the absense of a clear-cut relevance effect might be due to the difficulty manipulating this dimension contextually. It might be that the vignettes used in the experiment failed to make the listener relevance salient enough for any effects to be borne out. Further experiments should improve this manipulation, as well as additionally consider speaker relevance.

For the trigger ``or'', in contrast, we only find a positive effect of competence ($P =  0.993$, Fig.\ref{posteriors-main}, \texttt{competence\_xor} line, green color).  We find no comprehensive effects for relevance or competence. While the reason for the former might be the same as for ``some'', we suggest that these results might also be influenced by collinearity of the effects. We turn to this conjecture in light of the exploratory analyses, described below. 
Overall, these results would provide evidence against the scalar implicature hypothesis because not all factors yielded an effect on the robustness of scalar implicatures generated from ``some'' and ``or''. Similarly, the results would provide evidence against the identity hypothesis, as en effect of competence was observed for ``some'', but not for ``or''.

\subsection{Exploratory analysis}
\begin{figure*}[t]
	\begin{center}
		\includegraphics[width=1\linewidth]{images/posterior-effects-categorical.png}
	\end{center}
	\vspace{-0.3cm}
	\caption{Exploratory analysis with categorical predictors: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''. The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).}
	\label{posteriors-cat}
\end{figure*}
In addition to the main analysis, we conducted two exploratory analyses. First, we fit a model using categorical predictors only. That is, we fit a Bayesian linear mixed effects regression model, regressing the inference rating against our prior intuitive classifications of the vignettes on the dimensions relevance, competence and prior (high or low, respectively), the effect of trigger, and their interactions. The random effects structure and the priors were identical to the main model. The factors were sum-coded, -1 representing the low condition, 1 representing the high condition. 
This model decouples the predictors by dichotomizing their classification and, therefore, addresses the potential collinearty issue. Furthermore, it provides a comparison between participants' judgements and our intuitive prior classifications of the vignettes.

Given this model and the data, we found very similar results for both triggers: We observe a negative effect of prior and a positive effect of competence, in line with predictions of the scalar implicature hypothesis (Fig.\ref{posteriors-cat}). However, we again observe no effects of relevance. If anything, the relevance effect tends in the direction opposite to predictions of the SI account for the trigger ``or''. 
\begin{figure*}[t]
	\begin{center}
		\includegraphics[width=1\linewidth]{images/posterior-effects-means.png}
	\end{center}
	\vspace{-0.3cm}
	\caption{Exploratory analysis with predictor ratings averaged by-vignette: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''. The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).}
	\label{posteriors-mean}
\end{figure*}

These results match the second exploratory analysis, whereby the ratings were used as predictors in the linear model, but the ratings were averaged for each vignette across participants. In this second  exploratory model, we again find similar effects for ``some'' and ``or'': a positive effect of comptenece, negative for prior, and an indefinite relevance effect (Fig.\ref{posteriors-mean}). By averaging over single participant ratings, this model indicates that the absent competence effect in the main model might be due to by-participant variation.

In order to further investigate the relevance effects and potential collinarity concerns, we computed exploratory pairwise correlations of all the predictors. Interestingly, we find a significant correlation between the prior and the relevance effects for the trigger ``or'' ($R^2 = -0.106$). We also found a significant correlation between competence and relevance ($R^2 = 0.127$) for ``or''. No correlations were found for ``some''. These intriguing results might indicate that these effects are closely related for the disjunction. Intuitively, a negative correlation between prior and relevance of the stronger alternative ``and'' makes sense: the more likely the conjunction is a priori, the more likely it is that the listener might know about it already, yielding the implicature irrelevant to the listener. Similarly, competence might be connected to relevance indirectly via the prior: the more likely the stronger alternative is a priori, the more likely it is that a speaker knows about it, while the irrelevance still persists for the listener. However, this entangled results need further research.

\section{Discussion}
In sum, the results provide moderate evidence for the scalar implicature account of ``some'' and ``or''. While we could confirm the effects of prior and comptence for the interpretation of ``some'' as predicted by the SI hypothesis, we could not verify them for the disjunction. We also did not find evidence for effects of relevance on the robustness of scalar implicature generation, given the vignettes. Furthermore, our results hint at intriguing dependencies between relevance and prior effects for the interpretation of exclusive disjunctions. Taken together, the results might speak against the identity hypothesis, indicating that there might be subtle differences in the conceptualization of the stronger alternative for ``some'' and ``or''.

\section{Appendix}
\subsection{Vignettes}
Example transcript of an “or”-story and the associated statements:

Story: “Bill and Eric went to a bar. Eric’s wife, Lily, promised to pick them up if Eric drank any alcohol. She phones up Bill to ask about their night.”
Example comprehension statements: “Eric and Lily went to a bar together.” (false), “Eric and Bill went to see a football game.” (false), “Eric and Bill go out to a bar every weekend.” (uncertain), “Eric's wife has a driver’s licence.” (true)
Competence statement: “Bill knows whether Eric drank both wine and vodka.”
Relevance statement: “It is important for Lily to know whether her husband drank both wine and vodka.”
Prior statements: “If Eric drank wine, it is likely that he drank vodka as well.”, “If Eric drank vodka, it is likely that he drank wine as well.”
Critical utterance: “Bill tells Lily: 'Your husband drank wine or vodka.'”
Implicature statement: “From what Bill said we may conclude that Eric did not drink both wine and vodka.”

Example transcript of a “some”-story and the associated statements:

Story: “Bernard just finished his final exam. He isn't the brightest student, but if he passes this one, he will be through to the next grade. During the break, he goes to the teacher's lounge. He cannot wait to ask his teacher whether he passed. Bernard's teacher only just started correcting his exam.”
Example comprehension statements: “Bernard wants to get to the next grade.” (true), “Bernard wants to know his grade for his college application.” (false), “Bernard took a final exam in biology.” (uncertain), “Bernard is not the best student of his class.” (true)
Competence statement: “Bernard’s teacher knows whether all of the answers Bernard gave on the exam are correct.”
Prior statement: “If at least some of the answers Bernard gave on the exam are correct, then all of them are correct.”
Relevance statement: “It is important for Bernard to know whether all of the answers he gave on the exam are correct.”
Critical utterance: “Bernard’s teacher says: ‘Some of the answers you gave are correct.’”
Implicature statement: “From what Bernard’s teacher said we may conclude that not all of Bernard’s answers are correct.”

\bibliography{paper.bib}

\begin{addresses}
  \begin{address}
    Author1 Full Name \\
    Street \\
    City, State Zip \ldots \\
    \email{author1@example.org}
  \end{address}
  % Repeat or remove additional addresses as needed.
  \begin{address}
    Author2 Full Name \\
    Street \\
    City, State Zip \dots \\
    \email{author2@example.com}
  \end{address}
\end{addresses}

\end{document}