### 2021-12-20 analysis/main/05_xor-some_analysis_model_compariso_grid-search-SD.Rmd
full SD grid search vs Int

---
title: "Xor-Some Preregistered Study: Analyses"
author: "Polina Tsvilodub"
date: "8/31/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(brms)
library(tidyboot)
library(tidybayes)
```

Read data created in preprocessing script:
```{r, message=FALSE, warning=FALSE, results='hide'}
d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")

d_critical_zScored_wide_xor_prior_order <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-order.csv")

d_critical_zScored_wide_xor_prior_type <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-type.csv")

d_critical_zScored_wide_xor <- d_critical_zScored_wide %>% filter(main_type == "xor")
```

## Comparison of prior and relevance effects

Since we observed potential colinearity effects in the main preregistered analysis for the trigger "or", we further explore the effects of this colinearity on the strength of the individual effects in the regression models.

To do so, we compute individual models omitting one of the factors on the "or" data, and compute the Bayes Factor (BF) of the model omitting one of the colinear predictors, relative to the full model.

Upon the [first comparison](https://github.com/magpie-ea/magpie-xor-experiment/blob/master/analysis/main/04_xor-some_analysis_model_comparison.md) of different models, we decide to focus on pitting the addition of the relevance effect against the addition of the prior effect to the competence effect. For more stable results, we conduct a little grid search over the SD parameter of the prior for the main effect regression coefficients, since the BF might be quite sensitive to a particular configuration. The range of the SD parameter is motivated by the theoretically informed expected effect size.

```{r}
# define a convenience function for iterating over the SDs and critical models
models_grid_search_SD = function(sd_vec, iter) {
  # create output df
  results <- tribble(
    ~"model in favor", ~"model against", ~"SD", ~"BF"
  )
  # set priors
  for (i in sd_vec) {
    dist <- paste("student_t(1, 0, ", i, ")", sep = "")
    prior <- set_prior(dist, class = "b")
    # fit rel * comp model with specific prior
    model_rel_comp <- brm(target ~ relevance*competence +
                   (1 + relevance + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE
                 )
    # fit prior * comp model with specific prior
    model_prior_comp <- brm(target ~ prior*competence +
                   (1 + prior + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE
                 )
    # compute BF
    bf <- bayes_factor(model_rel_comp, model_prior_comp, silent = TRUE)
    print(paste("SD: ", i, sep=""))
    print(paste("BF: ", bf, sep=""))
    # append to results dataframe, write out
    results <- results %>% add_row(
      "model in favor" = "Rel * Comp",
      "model against" = "Prior * Comp",
      SD = i,
      BF = round(bf[1]$bf, 4)
    )
    # dot might be a problem
    write_csv(results, 
              paste("../../data/main/grid_rel-pri_SD_0_", i*100, ".csv", sep = ""))
  }
  return (results)
}
```

Call the grid search:
```{r, message=FALSE, warning=FALSE}
grid_results <- models_grid_search_SD(c(0.05, 0.1, 0.15), 40000)

```

```{r}
grid_results

write_csv(grid_results, "../../data/main/grid_rel-pri_SD_final.csv")
```

Plot the effect of SD against the BF:
```{r}
grid_results_final <- read_csv("../../data/main/grid_rel-pri_SD_final.csv")

grid_results_final %>%
ggplot(., aes(x = SD, y = BF, color = `model in favor`)) +
  geom_point() +
  geom_line() +
  ggtitle("BF in favor of including a relevance effect over\n including a prior effect, depending on SD of prior t-distribution")

ggsave("BF_vs_SD_rel-comp_pri-comp.pdf", width = 6, height = 4)
```

For exploration, we also compute the BF in favor of a model with an interaction between the two effects against the same model without interaction estimates.
```{r, results='hide', message=FALSE, warning=FALSE}
# use rather tight prior
priors <- set_prior("student_t(1, 0, 0.1)", class = "b")
# model with prior effect only
model_prior_comp_int <- brm(target ~ prior*competence +
                   (1 + prior + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )
model_prior_comp <- brm(target ~ prior + competence +
                   (1 + prior + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )
bf_pri_comp_int <- bayes_factor(model_prior_comp_int, model_prior_comp, silent = TRUE)
```

```{r}
summary(model_prior_comp_int)
summary(model_prior_comp)
bf_pri_comp_int
```

```{r}
model_rel_comp_int <- brm(target ~ relevance*competence +
                   (1 + relevance + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )

model_rel_comp <- brm(target ~ relevance + competence +
                   (1 + relevance + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )
bf_rel_comp_int <- bayes_factor(model_rel_comp_int, model_rel_comp, silent = TRUE)
```

```{r}
summary(model_rel_comp)
summary(model_rel_comp_int)
bf_rel_comp_int
```

For purpose of exploration, we also compare the full model with all interactions to a model without interactions, both with rather tight priors:
```{r, results='hide', message=FALSE, warning=FALSE}
# Full model 
# set priors
priors <- set_prior("student_t(1, 0, 0.1)", class = "b")

model_SI_ints <- brm(target ~ prior*competence*relevance +
                   (1 + prior + competence + relevance || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.95),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                )

model_SI <- brm(target ~ prior + competence + relevance +
                   (1 + prior + competence + relevance || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.95),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                )

bf_ints <- bayes_factor(model_SI_ints, model_SI, silent = TRUE)
```

```{r}
summary(model_SI)
summary(model_SI_ints)
bf_ints
```

Export exploration results to csv:
```{r}
results_ints <- tribble(
  ~"Model in favor", ~"Model against", ~"BF", ~"SD",
  "Pri * Comp", "Pri + Comp", bf_pri_comp_int[1]$bf, "0.1",
  "Rel * Comp", "Rel + Comp", bf_rel_comp_int[1]$bf, "0.1",
  "Rel * Pri * Comp", "Rel + Pri + Comp", bf_ints[1]$bf, "0.1"
)
results_ints 
results_ints %>% write_csv(., "../../data/main/bf_interactions_comparisons.csv")
```

## Model comparisons with grid search for final paper
Below, we conduct a similar grid search over more models. More specifically, we compute the BF in favor of a more complex model over the intercept-only baseline model. We compute these pairwise comparisons for each iteratively added explaantory factor. This results in 7 model comparisons. For each comparison, we perform a small grid search over the SD parameter of the prior t-distribution over the fixed effects regression coefficients, using the values [0.05, 0.1, 0.15]. For each model, 40.000 samples are extracted.
```{r}
# define a convenience function for iterating over the SDs and critical models
models_grid_search_SD_full = function(sd_vec, iter, trigger) {
  # create output df
  results <- tribble(
    ~"model in favor", ~"model against", "~trigger", ~"SD", ~"BF"
  )
  # filter data by trigger
  data <- d_critical_zScored_wide %>% filter(main_type == trigger)
  # set priors
  for (i in sd_vec) {
    dist <- paste("student_t(1, 0, ", i, ")", sep = "")
    prior <- set_prior(dist, class = "b")
    # fit the baseline intercept only model with same prior SD
    prior_int <- set_prior(dist, class = "Intercept")
    model_int <- brm(target ~ 1 +
                   (1 | submission_id) +
                   (1 | title),
                 data = data,
                 prior = prior_int,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE
                 )
    # fit the single predictor models
    model_pri <- brm(target ~ prior + 
                   (1 + prior || submission_id) +
                   (1 | title),
                 data = data,
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE
                 )
    model_rel <- brm(target ~ relevance + 
                   (1 + relevance || submission_id) +
                   (1 | title),
                 data = data,
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE
                 )
    model_comp <- brm(target ~ competence + 
                   (1 + competence || submission_id) +
                   (1 | title),
                 data = data,
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE
                 )
    # fit rel * comp model with specific prior
    model_rel_comp <- brm(target ~ relevance*competence +
                   (1 + relevance + competence || submission_id) +
                   (1 | title),
                 data = data,
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE
                 )
    # fit prior * comp model with specific prior
    model_pri_comp <- brm(target ~ prior*competence +
                   (1 + prior + competence || submission_id) +
                   (1 | title),
                 data = data,
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE
                 )
    # rel*pri model
    model_rel_pri <- brm(target ~ prior*relevance +
                   (1 + prior + relevance || submission_id) +
                   (1 | title),
                 data = data,
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE
                 )
    # fit full model
    model_pri_rel_comp <- brm(target ~ prior*competence*relevance +
                   (1 + prior + competence + relevance || submission_id) +
                   (1 | title),
                 data = data,
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE
                 )
    # compute BF
    # iterate over list of models 
    model_list <- c("model_rel", "model_pri", "model_comp", "model_rel_comp", "model_pri_comp", "model_rel_pri", "model_pri_rel_comp")
    for (m in model_list) {
      bf <- bayes_factor(m, model_int, silent = TRUE)
      print(paste("SD: ", i, sep=""))
      print(paste("Model in favor: ", m, sep=""))
      print(paste("BF: ", bf, sep=""))
      # append to results dataframe, write out
      results <- results %>% add_row(
        "model in favor" = m,
        "model against" = "Int",
        "trigger" = trigger,
        SD = i,
        BF = round(bf[1]$bf, 4)
      )
    write_csv(results, 
              paste("../../data/main/grid_full_", trigger, "_SD_0_", i*100, ".csv", sep = ""))
    }

  }
  return (results)
}
```

```{r}
# call the grid search for xor 
grid_results_full_xor <- models_grid_search_SD_full(c(0.05, 0.1, 0.15), 40000, "xor")

# call the grid search for some 
grid_results_full_some <- models_grid_search_SD_full(c(0.05, 0.1, 0.15), 40000, "some")

# write out results
write_csv(grid_results_full_xor, "../../data/main/grid_full_xor_SD_final.csv")
write_csv(grid_results_full_some, "../../data/main/grid_full_some_SD_final.csv")

```

Plot the effect of SD against the BF:
```{r}
grid_results_full_xor <- read_csv("../../data/main/grid_full_xor_SD_final.csv")
grid_results_full_some <- read_csv("../../data/main/grid_full_some_SD_final.csv")
grid_results_both <- rbind(grid_results_full_xor, grid_results_full_some)

grid_results_both %>%
ggplot(., aes(x = SD, y = BF, color = `model in favor`)) +
  geom_point() +
  geom_line() +
  facet_wrap(~trigger) + 
  ggtitle("BF in favor of including a relevance effect over\n including a prior effect, depending on SD of prior t-distribution")

#ggsave("BF_vs_SD_all_vs_int.pdf", width = 8, height = 4)
```

```{r}

# bf <- bayes_factor(model_rel_comp, model_prior_comp, silent = TRUE)
#     print(paste("SD: ", i, sep=""))
#     print(paste("BF: ", bf, sep=""))
#     # append to results dataframe, write out
#     results <- results %>% add_row(
#       "model in favor" = "Rel * Comp",
#       "model against" = "Prior * Comp",
#       SD = i,
#       BF = round(bf[1]$bf, 4)
#     )
#     # dot might be a problem
#     write_csv(results, 
#               paste("../../data/main/grid_rel-pri_SD_0_", i*100, ".csv", sep = ""))
```