### 2022-05-12 analysis/main/mf-exploration.Rmd
added exploration files (role of by-item REs)

---
title: "Exploring influence of REs on models fits"
author: "Michael Franke"
date: "12/05/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = F, message = FALSE, warning = FALSE, error = FALSE, fig.width = 5, fig.align = "center")
```

```{r libraries, message = FALSE, warning = FALSE, include = FALSE}
library(tidyverse)
library(brms)
library(tidyboot)
library(tidybayes)
```

# Data

We use the data with the "new" z-scoring for the prior data for 'or'.

```{r}
d_critical_clean <- read_csv("../../data/main/results_prereg_tidy_noDuplicates_zScored_meansFirst_wide.csv") %>% 
  mutate(main_type = factor(main_type))
head(d_critical_clean)
```

We also want additional information about the vignettes, i.e., their intuitive pre-classifications.

```{r }
vignettes <- read_csv("./../../data/main/results_73_xor-some-Prolific-main_N275_anonym.csv") %>% 
  select(title, main_type, relevance, competence, prior) %>% 
  unique() %>% 
  filter(! is.na(main_type)) %>% 
  rename(relCat = relevance,
         compCat = competence,
         priCat = prior)
head(vignettes)
```

# Functions for fitting models

We are interested in comparing the following models, specified here in terms of `brms` formulas:

```{r }
## full model with REs for subjects and items
f_full <- formula(
  target ~ prior * competence * relevance * main_type +
    (1 + prior + competence + relevance + main_type || submission_id) +
    (1 | title)
)

## model with RE for 'title' only (by-item random intercepts)
f_RETitle <- formula(
  target ~ prior * competence * relevance * main_type +
    (1 | title)
)

## RE for 'submject_id' only (by-subj random intercepts)
f_RESubj <- formula(
  target ~ prior * competence * relevance * main_type +
    (1 + prior + competence + relevance + main_type || submission_id)
)

## model without REs; fixed effects only 
f_noREs <- formula(
  target ~ prior * competence * relevance * main_type
)
```

Here's a function for running the models:

```{r }

run_model <- function(form, data){

  ## some = 0, xor = 1
  contrasts(data$main_type)

  ## set priors
  priors <- set_prior("student_t(1, 0, 2)", class = "b")

  ## run and return model
  brm(
    formula      = form,
    data         = data,
    prior        = priors,
    sample_prior = T,
    control      = list(adapt_delta = 0.95),
    cores        = 4,
    iter         = 3000
  )

} 
```

And here is a function for returning the relevant result summary from a fitted model:

```{r }
## function to get prediction for a model fit
get_predictions <- function(model_SI) {
  model_SI %>% spread_draws(b_Intercept, b_prior, b_competence, b_relevance,
                            b_main_typexor, `b_prior:competence`,
                            `b_prior:relevance`, `b_competence:relevance`,
                            `b_prior:main_typexor`, `b_competence:main_typexor`,
                            `b_relevance:main_typexor`, `b_prior:competence:relevance`, 
                            `b_prior:competence:main_typexor`, `b_prior:relevance:main_typexor`,
                            `b_competence:relevance:main_typexor`, 
                            `b_prior:competence:relevance:main_typexor`) %>% 
    mutate(
      prior_xor = b_prior + `b_prior:main_typexor`,
      prior_some = b_prior,
      competence_xor = b_competence + `b_competence:main_typexor`,
      competence_some = b_competence,
      relevance_xor =  b_relevance + `b_relevance:main_typexor`,
      relevance_some = b_relevance
    ) -> model_SI_posteriors

  # check P that the effects are positive / negative / no effect present
  posterior_hypotheses <- model_SI_posteriors %>% 
    select(prior_xor, prior_some, 
           competence_xor, competence_some,
           relevance_xor, relevance_some) %>%
    gather(key, val) %>%
    group_by(key) %>% mutate(positive = mean(val > 0.05),
                             negative = mean(val < -0.05),
                             no = mean(val %>% between(-0.05, 0.05))) %>%
    summarise(positive_eff = mean(positive),
              no_eff = mean(no),
              negative_eff = mean(negative))
  posterior_hypotheses
}
```

# Fitting and inspecting the relevant models

Model fits:

```{r results = 'hide'}
model_SI         <- run_model(f_full,d_critical_clean) 
model_SI_RETitle <- run_model(f_RETitle,d_critical_clean)
model_SI_RESubj  <- run_model(f_RESubj,d_critical_clean)
model_SI_noREs   <- run_model(f_noREs,d_critical_clean)
```

We can now inspect the results for each model, and find that **including by-item random intercepts determines whether we see a main effect of prior for the 'or' items, irrespective of whether we include or omit the by-subject random effects**.
We may conclude from this that it is really the by-item random effects that "explain away" the main effect of prior on 'or' items.

```{r }
## We compare the results for three model which differ only in the random effects
## they take into account. We care about a potential main effect of 'prior' in the
## 'or' condition.

## (i)  full REs (subj. & item)  -> no effect 
get_predictions(model_SI)

## (ii) no subj., just item REs  -> no effect 
get_predictions(model_SI_RETitle)

## (iii) just subj., no item REs -> strong effect 
get_predictions(model_SI_RESubj)

## (iv) no REs                   -> strong effect 
get_predictions(model_SI_noREs)
```

# Probing which vignettes are most influential in producing/hiding the effect

Which vignettes does the model estimate large random by-item effect for?
Here's a function that prints the top 6 most extreme estimates for the random effects in question.

```{r}

extract_REs <- function(model_fit) {
  randEff <- brms::ranef(model_fit)$title %>% as.data.frame() %>% 
    as_tibble(rownames = "title")
  vignettes  %>% 
    full_join(randEff, by= 'title')
}

```

Applying this function to both model with by-item random effects, we see that these top six do not differ much between whether we include the by-subject random effects or not.

```{r}
## full model
extract_REs(model_SI) %>% arrange(Estimate.Intercept)
extract_REs(model_SI) %>% arrange(-Estimate.Intercept)

## model without by-subject REs
extract_REs(model_SI_RETitle) %>% arrange(Estimate.Intercept)
extract_REs(model_SI_RETitle) %>% arrange(-Estimate.Intercept)

```

To further find out how the addition of by-item REs affects the model fit, we can try fittin a model *with* by-item REs, but omitting vignettes which receive rather high/low estimates for their RE.
Here we gather the titles of the top-$n$ highest or lowest ranked vignettes for each trigger word.

```{r}
get_extreme_vignettes <- function(model = model_SI, n=5, trigger_type = 'some', high = T) {
  extract_REs(model) %>% arrange(Estimate.Intercept * ifelse(high, 1, -1)) %>% 
    filter(main_type == trigger_type) %>% 
    pull(title) %>% .[1:n]
}

hi_some <- get_extreme_vignettes(trigger_type = 'some', high = T)
lo_some <- get_extreme_vignettes(trigger_type = 'some', high = F)
hi_or   <- get_extreme_vignettes(trigger_type = 'xor', high = T)
lo_or   <- get_extreme_vignettes(trigger_type = 'xor', high = F)
```

So, let's fit some model on reduced data sets, excluding "extreme vignettes".

```{r, results = 'hide'}

predictions_for_reduced_data <- function(excluded_vignettes) {
  run_model(
    f_full, 
    d_critical_clean %>% filter(! title %in% excluded_vignettes)
  ) %>% get_predictions()  
}

predictions_for_reduced_data(hi_some)
predictions_for_reduced_data(lo_some)
predictions_for_reduced_data(hi_or)
predictions_for_reduced_data(lo_or)

```

It seems that the vignettes for 'some' which are estimated to have particularly high random intercepts are most conducive for whether the main effect shows or not.