### 2022-05-13 analysis/main/mf-exploration.Rmd
add joint posterior plots

---
title: "Exploring influence of REs on models fits"
author: "Michael Franke"
date: "12/05/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = F, message = FALSE, warning = FALSE, error = FALSE, fig.width = 5, fig.align = "center")
```

```{r libraries, message = FALSE, warning = FALSE, include = FALSE}
library(tidyverse)
library(brms)
library(tidyboot)
library(tidybayes)
```

# Data

We use the data with the "new" z-scoring for the prior data for 'or'.

```{r}
d_critical_clean <- read_csv("../../data/main/results_prereg_tidy_noDuplicates_zScored_meansFirst_wide.csv") %>% 
  mutate(main_type = factor(main_type))
head(d_critical_clean)
```

We also want additional information about the vignettes, i.e., their intuitive pre-classifications.

```{r }
vignettes <- read_csv("./../../data/main/results_73_xor-some-Prolific-main_N275_anonym.csv") %>% 
  select(title, main_type, relevance, competence, prior) %>% 
  unique() %>% 
  filter(! is.na(main_type)) %>% 
  rename(relCat = relevance,
         compCat = competence,
         priCat = prior)
head(vignettes)
```

# Functions for fitting models

We are interested in comparing the following models, specified here in terms of `brms` formulas:

```{r }
## full model with REs for subjects and items
f_full <- formula(
  target ~ prior * competence * relevance * main_type +
    (1 + prior + competence + relevance + main_type || submission_id) +
    (1 | title)
)

## model with RE for 'title' only (by-item random intercepts)
f_RETitle <- formula(
  target ~ prior * competence * relevance * main_type +
    (1 | title)
)

## RE for 'submject_id' only (by-subj random intercepts)
f_RESubj <- formula(
  target ~ prior * competence * relevance * main_type +
    (1 + prior + competence + relevance + main_type || submission_id)
)

## model without REs; fixed effects only 
f_noREs <- formula(
  target ~ prior * competence * relevance * main_type
)

## full model with RE ineractions by-subject
f_full_inters <- formula(
  target ~ prior * competence * relevance * main_type +
    (1 + prior * competence * relevance * main_type || submission_id) +
    (1 | title)
)
```

Here's a function for running the models:

```{r }

run_model <- function(form, data){

  ## some = 0, xor = 1
  contrasts(data$main_type)

  ## set priors
  priors <- set_prior("student_t(1, 0, 2)", class = "b")

  ## run and return model
  brm(
    formula      = form,
    data         = data,
    prior        = priors,
    sample_prior = T,
    control      = list(adapt_delta = 0.95),
    cores        = 4,
    iter         = 3000
  )

} 
```

And here is a function for returning the relevant result summary from a fitted model:

```{r }
## function to get prediction for a model fit
get_predictions <- function(model_SI) {
  model_SI %>% spread_draws(b_Intercept, b_prior, b_competence, b_relevance,
                            b_main_typexor, `b_prior:competence`,
                            `b_prior:relevance`, `b_competence:relevance`,
                            `b_prior:main_typexor`, `b_competence:main_typexor`,
                            `b_relevance:main_typexor`, `b_prior:competence:relevance`, 
                            `b_prior:competence:main_typexor`, `b_prior:relevance:main_typexor`,
                            `b_competence:relevance:main_typexor`, 
                            `b_prior:competence:relevance:main_typexor`) %>% 
    mutate(
      prior_xor = b_prior + `b_prior:main_typexor`,
      prior_some = b_prior,
      competence_xor = b_competence + `b_competence:main_typexor`,
      competence_some = b_competence,
      relevance_xor =  b_relevance + `b_relevance:main_typexor`,
      relevance_some = b_relevance
    ) -> model_SI_posteriors

  # check P that the effects are positive / negative / no effect present
  posterior_hypotheses <- model_SI_posteriors %>% 
    select(prior_xor, prior_some, 
           competence_xor, competence_some,
           relevance_xor, relevance_some) %>%
    gather(key, val) %>%
    group_by(key) %>% mutate(positive = mean(val > 0.05),
                             negative = mean(val < -0.05),
                             no = mean(val %>% between(-0.05, 0.05))) %>%
    summarise(positive_eff = mean(positive),
              no_eff = mean(no),
              negative_eff = mean(negative))
  posterior_hypotheses
}
```

# Fitting and inspecting the relevant models

Model fits:

```{r results = 'hide'}
model_SI         <- run_model(f_full,d_critical_clean) 
model_SI_RETitle <- run_model(f_RETitle,d_critical_clean)
model_SI_RESubj  <- run_model(f_RESubj,d_critical_clean)
model_SI_noREs   <- run_model(f_noREs,d_critical_clean)
model_SI_fullREs <- run_model(f_full_inters, d_critical_clean)
```

We can now inspect the results for each model, and find that **including by-item random intercepts determines whether we see a main effect of prior for the 'or' items, irrespective of whether we include or omit the by-subject random effects**.
We may conclude from this that it is really the by-item random effects that "explain away" the main effect of prior on 'or' items.

```{r }
## We compare the results for three model which differ only in the random effects
## they take into account. We care about a potential main effect of 'prior' in the
## 'or' condition.

## (i)  full REs (subj. & item)  -> no effect 
get_predictions(model_SI)

## (ii) no subj., just item REs  -> no effect 
get_predictions(model_SI_RETitle)

## (iii) just subj., no item REs -> strong effect 
get_predictions(model_SI_RESubj)

## (iv) no REs                   -> strong effect 
get_predictions(model_SI_noREs)
```

# Probing which vignettes are most influential in producing/hiding the effect

Which vignettes does the model estimate large random by-item effect for?
Here's a function that prints the top 6 most extreme estimates for the random effects in question.

```{r}

extract_REs <- function(model_fit) {
  randEff <- brms::ranef(model_fit)$title %>% as.data.frame() %>% 
    as_tibble(rownames = "title")
  vignettes  %>% 
    full_join(randEff, by= 'title')
}

```

Applying this function to both model with by-item random effects, we see that these top six do not differ much between whether we include the by-subject random effects or not.

```{r}
## full model
extract_REs(model_SI) %>% arrange(Estimate.Intercept)
extract_REs(model_SI) %>% arrange(-Estimate.Intercept)

## model without by-subject REs
extract_REs(model_SI_RETitle) %>% arrange(Estimate.Intercept)
extract_REs(model_SI_RETitle) %>% arrange(-Estimate.Intercept)

```

To further find out how the addition of by-item REs affects the model fit, we can try fittin a model *with* by-item REs, but omitting vignettes which receive rather high/low estimates for their RE.
Here we gather the titles of the top-$n$ highest or lowest ranked vignettes for each trigger word.

```{r}
get_extreme_vignettes <- function(model = model_SI, n=5, trigger_type = 'some', high = T) {
  extract_REs(model) %>% arrange(Estimate.Intercept * ifelse(high, 1, -1)) %>% 
    filter(main_type == trigger_type) %>% 
    pull(title) %>% .[1:n]
}

hi_some <- get_extreme_vignettes(trigger_type = 'some', high = T)
lo_some <- get_extreme_vignettes(trigger_type = 'some', high = F)
hi_or   <- get_extreme_vignettes(trigger_type = 'xor', high = T)
lo_or   <- get_extreme_vignettes(trigger_type = 'xor', high = F)
```

So, let's fit some model on reduced data sets, excluding "extreme vignettes".

```{r, results = 'hide'}

predictions_for_reduced_data <- function(excluded_vignettes) {
  run_model(
    f_full, 
    d_critical_clean %>% filter(! title %in% excluded_vignettes)
  ) %>% get_predictions()  
}

predictions_for_reduced_data(hi_some)
predictions_for_reduced_data(lo_some)
predictions_for_reduced_data(hi_or)
predictions_for_reduced_data(lo_or)

```

It seems that the vignettes for 'some' which are estimated to have particularly high random intercepts are most conducive for whether the main effect shows or not.

## Polina's shenanigans

Check the joint posterior of fixed and random effects, to investigate which of them might cause the pattern of results that we observe. Analysis suggested by Roger.
The following plots shows the joint posterior of fixed and random effects in the xor and some conditions, respectively.

``` {r}
# extract posterior contrast of interest of fixed effects
posteriors <- model_SI_fullREs %>%
  spread_draws(b_prior, `b_prior:main_typexor`, b_relevance, `b_relevance:main_typexor`, b_competence, `b_competence:main_typexor`,
               sd_submission_id__prior, `sd_submission_id__prior:main_typexor`,
               sd_submission_id__Intercept, sd_title__Intercept, 
               sd_submission_id__relevance, `sd_submission_id__relevance:main_typexor`) %>%
    # extract contrasts of interest, especially effect of syntax by-trial
    mutate(prior_xor = b_prior + `b_prior:main_typexor`,
           prior_some = b_prior,
           relevance_xor = b_relevance + `b_relevance:main_typexor`,
           relevance_some = b_relevance,
           competence_xor = b_competence + `b_competence:main_typexor`,
           competence_some = b_competence,
           random_prior_xor = sd_submission_id__prior + `sd_submission_id__prior:main_typexor`,
           random_prior_some = sd_submission_id__prior,
           random_relevance_xor = sd_submission_id__relevance + `sd_submission_id__relevance:main_typexor`,
           random_relevance_some = sd_submission_id__relevance
           ) %>% # subject vs predicate syntax
    select(prior_xor, prior_some, relevance_xor, relevance_some, competence_xor, competence_some,
           random_prior_xor, random_prior_some, sd_submission_id__Intercept, sd_title__Intercept, random_relevance_xor, random_relevance_some) %>%
    gather(key, val)

```

Helper function for extracting and plotting a heat map of two effects, given a df containing all relevant posterior samples / effects, the effects and a title for the plot. 
```{r}
get_and_plot_2d_posterior <- function(posteriors, fe, re, title) {
  fe_selected <- posteriors %>% filter(key == fe) %>% rename(fe = val)
  re_selected <- posteriors %>% filter(key == re) %>% rename(re = val) %>% select(-key)

  df <- tibble(fe_selected, re_selected)

  p <- ggplot(df, aes(x = fe, y = re)) +
  stat_bin2d(aes(fill = after_stat(density))) +
  ggtitle(title)

  return (p)
}
```

First, we explore how the fixed effects of prior behave against by-title random intercepts and by-subject random slope effects of prior.
```{r}
get_and_plot_2d_posterior(posteriors, "prior_xor", "sd_title__Intercept", "FE of prior for XOR against by-story random intercepts")
#ggsave("FE_prior_RE_story_xor.pdf", width=7, height=5)

get_and_plot_2d_posterior(posteriors, "prior_some", "sd_title__Intercept", "FE of prior for SOME against by-story random intercepts")

get_and_plot_2d_posterior(posteriors, "prior_xor", "random_prior_xor", "FE of prior for XOR against by-subj prior RE")

get_and_plot_2d_posterior(posteriors, "prior_some", "random_prior_some", "FE of prior for SOME against by-subj prior RE")
```

Now, double check if there is a "credit assignment" problem for other factors.
```{r}
get_and_plot_2d_posterior(posteriors, "competence_xor", "sd_title__Intercept", "FE of competence for XOR against by-story random intercepts")

get_and_plot_2d_posterior(posteriors, "competence_some", "sd_title__Intercept", "FE of competence for SOME against by-story random intercepts")

get_and_plot_2d_posterior(posteriors, "relevance_xor", "sd_title__Intercept", "FE of relevance for XOR against by-story random intercepts")

get_and_plot_2d_posterior(posteriors, "relevance_some", "sd_title__Intercept", "FE of relevance for SOME against by-story random intercepts")
```

Explore an xor-only model.
```{r}
f_xor <- formula(
  target ~ prior * competence * relevance +
    (1 + prior + competence + relevance || submission_id) +
    (1 | title)
)
model_xor<- run_model(f_xor, d_critical_clean %>% filter(main_type == "xor")) 
summary(model_xor)
```

```{r}
# get the probability of a negative effect for or
model_xor %>% spread_draws(b_prior) %>%
  select(b_prior) %>%
  gather(key, val) %>%
  group_by(key) %>% mutate(positive = mean(val > 0.05),
                             negative = mean(val < -0.05),
                             no = mean(val %>% between(-0.05, 0.05))) %>%
    summarise(positive_eff = mean(positive),
              no_eff = mean(no),
              negative_eff = mean(negative))

```